### 2016-05-04 paper.tex
Initial version

\documentclass{sig-alternate-05-2015}
\usepackage[l2tabu,orthodox]{nag}
\usepackage{fixltx2e}
\usepackage[utf8x]{inputenc}       
\usepackage[british]{babel}        
\usepackage[babel=true]{microtype}
\usepackage{amsmath}
\usepackage[all]{onlyamsmath}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{upquote}
\usepackage{graphicx}
\usepackage{url}
\usepackage[caption=false]{subfig}
\usepackage{booktabs}
\usepackage{bytefield}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{color}
\usepackage{balance}

\newcommand{\todo}[1]{\textbf{\textcolor{red}{To do: #1}}}

%==================================================================================================
% The following information gets written into the PDF file information:
\pdfinfo{
  /Title        (...)
  /Author       (...)
  /Subject      (...)
  /Keywords     (..., ..., ...)
  /CreationDate (D:20150827110616+01'00')
  /ModDate      (D:20150827110616+01'00')
  /Creator      (LaTeX)
  /Producer     (pdfTeX)
}
%==================================================================================================
\begin{document}
\title{Implementing Real-time Transport Services over an Ossified Network}
\numberofauthors{3}
\author{
  \alignauthor
    Stephen McQuistin\\
    \affaddr{University of Glasgow, UK}\\
    \email{sm@smcquistin.uk}
  \alignauthor
    Colin Perkins\\
    \affaddr{University of Glasgow, UK}\\
    \email{csp@csperkins.org}
  \alignauthor
    Marwan Fayed\\
    \affaddr{University of Stirling, UK}\\
    \email{mmf@cs.stir.ac.uk}
}

\toappear{}
\maketitle
%==================================================================================================
\begin{abstract}

% Four sentences:
%  - State the problem
%  - Say why it's an interesting problem
%  - Say what your solution achieves
%  - Say what follows from your solution

\end{abstract}
%==================================================================================================
\begin{CCSXML}
  <ccs2012>
    <concept>
      <concept_id>10003033.10003039.10003040</concept_id>
      <concept_desc>Networks~Network protocol design</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
    <concept>
      <concept_id>10003033.10003039.10003048</concept_id>
      <concept_desc>Networks~Transport protocols</concept_desc>
      <concept_significance>500</concept_significance>
    </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Networks~Network protocol design}
\ccsdesc[500]{Networks~Transport protocols}

\printccsdesc

\keywords{Transport protocols; real-time multimedia applications}
%==================================================================================================
\section{Introduction}

% General advice:
%  - Avoid stock and cliche phrases such as "recent advances in XYZ" or
%    anything alluding to the growth of the Internet (unless your paper
%    is about the growth of the Internet!).
%  - Be sure that the introduction lets the reader know what this paper is
%    about, not just how important your general area of research is.
%    Readers won't stick with you for three pages to find out what you are
%    talking about.
%  - The introduction must motivate your work by pinpointing the problem
%    you are addressing and then give an overview of your approach and/or
%    contributions (and perhaps even a general description of your
%    results).  In this way, the intro sets up my expectations for the
%    rest of your paper -- it provides the context, and a preview.
%  - Repeating the abstract in the introduction is a waste of space.

% para. 1: motivation: broadly, what is problem area, why important?

% para. 2: narrow down: what is problem you specifically consider

% para. 3: "In this paper, we ...": most crucial paragraph, tell your
% elevator pitch. The story is not what you did, but rather:
%  - what you show, new ideas, new insights
%  - why interesting, important?
% State your contributions: these drive the entire paper.  Contributions
% should be refutable claims, not vague generic statements.

In this paper, we ...

% para. 4: how different/better/relates to other work

% para. 5: "We structure the remainder of this paper as follows."

We structure the remainder of this paper as follows.

%==================================================================================================
% Concentrate single-mindedly on a narrative that:
%  - Describes the problem, and why it's interesting
%  - Describes your idea
%  - Defends your idea, showing how it solves the problem, and filling out
%    the details
% On the way, cite relevant work in passing, but defer discussion to the
% end.
%
% Introduce the problem, and your idea, using examples, and only then
% present the general case. Explain the idea as if your were speaking to
% someone using a whiteboard. Conveying the intuition is primary; details
% follow. Write in a top down manner: state broad themes and ideas first,
% then go into details.
%
% The introduction makes claims. The body of the paper provides evidence
% to support each claim. Check each claim in the introduction, identify
% the evidence, and forward-reference it from the claim. 
%==================================================================================================
\section{}

% The problem

%==================================================================================================
\section{}

% My idea

%==================================================================================================
\section{}

% The details

% Describe results carefully:
%  - clearly state assumptions
%  - give enough information to allow the reader to recreate the results
%  - ensure results are representative; statistically meaningful, etc.
%  - don't overstate results
%  - equally, don't understate them: consider the broader implications

%==================================================================================================
\section{Related Work}

% This should come near the end, and focussing on discussing how your work
% relates to that of others. Any relevant related work should have been
% cited already, so this is not a list of related work, it's a discussion
% of how that work relates.
%
% Why not put related work after the introduction? 1) because describing
% alternative approaches gets between the reader and your idea; and 2)
% because the reader knows nothing about the problem yet, so your
% (carefully trimmed) description of various technical trade-offs is
% absolutely incomprehensible.
% 
% When writing the related work:
%  - Give credit to others where it's due; this doesn't diminish the
%    credit you get from your paper. 
%  - Acknowledge weaknesses in your approach.
%  - Ensure related work is accurate and up-to-date

%==================================================================================================
\section{Conclusions}

%==================================================================================================
\section{Acknowledgements}

% Acknowledge funding sources.

%==================================================================================================
\bibliographystyle{abbrv}
\bibliography{paper}
\end{document}
% vim: set ts=2 sw=2 tw=75 et ai:

### 2016-05-04 bin/check-for-todo.sh

@@ -1,0 +1,9 @@
+#!/bin/sh
+
+if [ "$1" = "" ]; then
+ echo "usage: bin/check-for-todo.sh <file> ..."
+ exit
+fi
+ 
+egrep -H -i -n --color "\\\\todo\\{" $* | grep -v ".*%.*\\\\todo" || true
+ 

### 2016-05-04 bin/check-pdf-fonts.sh

@@ -1,0 +1,21 @@
+#!/bin/sh
+
+pdffonts $1 > $2
+
+nmf=`cat $2 | tail -n +3 | awk '{if ($(NF-4) != "yes") print $0}' | wc -l`
+
+if [ $nmf -gt 0 ]; then \
+  tput setaf 1 
+  tput bold
+  echo ""
+  echo "WARNING: the following fonts are not embedded:"
+  echo ""
+  tput sgr0
+  cat $2
+  echo ""
+  echo "Try running \"updmap --edit\" and setting \"pdftexDownloadBase14 true\""
+  echo ""
+  echo "================================================================================"
+fi
+
+

### 2016-05-04 bin/latex-build.sh

@@ -1,0 +1,137 @@
+#!/bin/sh
+
+if [ $# != 1 ]; then
+  echo "Usage: latex-build.sh <basename>"
+  exit 1
+fi
+
+TEX_BASE=$1
+
+REGEX_CITE="LaTeX Warning: Citation.*undefined"
+REGEX_LABL="LaTeX Warning: Label(s) may have changed. Rerun to get cross-references right."
+REGEX_BOOK="Package rerunfilecheck Warning: File .*out. has changed"
+
+COLS=`tput cols`
+BLANK_LINE=""
+
+# Set BLANK_LINE to be a line that matches the full width of the terminal
+i=0
+while [ $i -lt $COLS ]; 
+do
+  BLANK_LINE="$BLANK_LINE="
+  i=`expr $i + 1`
+done
+
+blank_line () {
+  tput setaf 2
+  echo $BLANK_LINE
+  tput sgr0
+}
+
+done_bib=0
+do_bib=0
+do_tex=1
+
+while [ $do_tex = 1 ]; do
+  blank_line
+
+  pdflatex -halt-on-error $TEX_BASE.tex
+  if [ $? = 1 ]; then
+    exit
+  fi
+  do_tex=0
+
+  # Rerun LaTeX if the labels have changed
+  labl_changed=`grep -c "$REGEX_LABL" $TEX_BASE.log`
+  if [ $labl_changed != 0 ]; then
+    do_tex=1
+  fi
+
+  # Rerun LaTeX if PDF bookmarks have changed
+  book_changed=`grep -c "$REGEX_BOOK" $TEX_BASE.log`
+  if [ $book_changed != 0 ]; then
+    do_tex=1
+  fi
+
+  # Are there are undefined citations?
+  undef_cite=`grep -c "$REGEX_CITE" $TEX_BASE.log`
+  if [ $undef_cite != 0 ]; then
+    if [ $done_bib == 0 ]; then 
+      do_bib=1
+    fi
+    if [ $done_bib == 1 ]; then
+      done_bib=2
+      do_tex=1
+    fi
+  fi
+
+  # Check if any of the *.bib files includes have been modified since
+  # BibTeX was last run; if so, request a new run of BibTeX
+  for f in `grep '\\\\bibdata{' $TEX_BASE.aux | sed 's/\\\bibdata{//' | sed 's/}//'`
+  do
+    if [ $f.bib -nt $TEX_BASE.bbl ]; then
+      do_bib=1
+    fi
+  done
+
[…]

### 2016-05-06 bin/latex-build.sh
Update pre-processing

@@ -95,36 +95,14 @@ blank_line
 # The first call to gs tries to embed all fonts. This should turn any
 # non-embedded fonts, of any type, into embedded Type 1 fonts, without
 # subsetting.
-echo "Post-processing PDF file... (1/3)"
+echo "Post-processing PDF file..."
 gs -q -dSAFER -dNOPAUSE -dBATCH -dCompatibilityLevel=1.4 -dDetectDuplicateImages=true \
    -dPDFSETTINGS=/prepress -dEmbedAllFonts=true -dSubsetFonts=false \
-   -sDEVICE=pdfwrite -sOutputFile=$TEX_BASE.tmp-a.pdf \
+   -sDEVICE=pdfwrite -sOutputFile=$TEX_BASE.tmp.pdf \
    -f $TEX_BASE.pdf
 
-# The second call to gs tries to un-embed any fonts in the PDF. If any
-# included image files have unwanted embedded fonts in TrueType format, 
-# this will turn them into non-embedded TrueType fonts.
-echo "Post-processing PDF file... (2/3)"
-gs -q -dSAFER -dNOPAUSE -dBATCH -dCompatibilityLevel=1.4 -dDetectDuplicateImages=true \
-   -dPDFSETTINGS=/prepress -dEmbedAllFonts=false -dSubsetFonts=false \
-   -sDEVICE=pdfwrite -sOutputFile=$TEX_BASE.tmp-b.pdf \
-   -c ".setpdfwrite <</AlwaysEmbed [ ]>> setdistillerparams" \
-   -f $TEX_BASE.tmp-a.pdf
-
-# The third call to gs tries to embed all fonts. This should turn any
-# non-embedded fonts, of any type, into embedded Type 1 fonts, and will
-# subset those fonts to include only the used characters, reducing the
-# size of the file.
-echo "Post-processing PDF file... (3/3)"
-gs -q -dSAFER -dNOPAUSE -dBATCH -dCompatibilityLevel=1.4 -dDetectDuplicateImages=true \
-   -dPDFSETTINGS=/prepress -dEmbedAllFonts=true -dSubsetFonts=true \
-   -sDEVICE=pdfwrite -sOutputFile=$TEX_BASE.tmp-c.pdf \
-   -f $TEX_BASE.tmp-b.pdf
-
[…]

### 2016-05-06 paper.tex
Update

@@
 \usepackage[l2tabu,orthodox]{nag}
-\usepackage{fixltx2e}
 \usepackage[utf8x]{inputenc}       
@@
 
-% General advice:
-%  - Avoid stock and cliche phrases such as "recent advances in XYZ" or
-%    anything alluding to the growth of the Internet (unless your paper
-%    is about the growth of the Internet!).
-%  - Be sure that the introduction lets the reader know what this paper is
-%    about, not just how important your general area of research is.
-%    Readers won't stick with you for three pages to find out what you are
-%    talking about.
-%  - The introduction must motivate your work by pinpointing the problem
-%    you are addressing and then give an overview of your approach and/or
-%    contributions (and perhaps even a general description of your
-%    results).  In this way, the intro sets up my expectations for the
-%    rest of your paper -- it provides the context, and a preview.
-%  - Repeating the abstract in the introduction is a waste of space.
+% A good paper introduction is fairly formulaic. If you follow a simple set
+% of rules, you can write a very good introduction. The following outline can
+% be varied. For example, you can use two paragraphs instead of one, or you
+% can place more emphasis on one aspect of the intro than another. But in all
+% cases, all of the points below need to be covered in an introduction, and
+% in most papers, you don't need to cover anything more in an introduction.
+%
+% Paragraph 1: Motivation. At a high level, what is the problem area you
+% are working in and why is it important? It is important to set the larger
+% context here. Why is the problem of interest and importance to the larger
+% community?
 
-% para. 1: motivation: broadly, what is problem area, why important?
+Real-time applications are increasingly important in the Internet. 
+We want to make it easier to write these applications, while improving
+the quality of experience for users of those applications by lowering
+latency and increasing robustness. 
+In this, we are challenged by the limitations of the standard Internet 
+transport protocols, and by the ossified nature of the network, which makes
+it increasingly difficult to deploy new transports.
 
-% para. 2: narrow down: what is problem you specifically consider
+% Paragraph 2: What is the specific problem considered in this paper? This
+% paragraph narrows down the topic area of the paper. In the first
+% paragraph you have established general context and importance. Here you
+% establish specific context and background.
 
-% para. 3: "In this paper, we ...": most crucial paragraph, tell your
-% elevator pitch. The story is not what you did, but rather:
+In practice, the only protocols that can be widely used in the Internet are
+TCP and UDP. Other protocols are blocked by firewalls and other middleboxes.  
+TCP provides sophisticated congestion control, coupled with a reliable,
+ordered, byte stream API. These are suitable for many applications, but are
+not necessarily appropriate for real-time traffic.  UDP exposes essentially
+the raw IP packet delivery service, offering flexibility to develop new
+protocols, but at the cost of requiring completely new protocol mechanisms
+to be defined. While both protocols are usable for real-time applications,
+neither really provides the right services and API.
+
+% Paragraph 3: "In this paper, we show that...". This is the key paragraph
+% in the introduction - you summarize, in one paragraph, what are the main
+% contributions of your paper, given the context established in paragraphs 
+% 1 and 2. What's the general approach taken? Why are the specific results
+% significant? The story is not what you did, but rather:
 %  - what you show, new ideas, new insights
@@
 
-In this paper, we ...
+In this paper we develop an appropriate set of transport services and APIs
+for real-time applications.
+Then, we demonstrate that it is possible to realise these services and APIs
+in the context of both TCP and UDP, despite the limitations imposed by the
+legacy protocols, middleboxes, and ossification of the network. 
+We present initial results to show that, despite the observed ossification,
+the network has the flexibility to deploy new transport protocols, if care
+is taken to reinterpret layer boundaries and deploy those new transports in
+the context of TCP and UDP.
 
-% para. 4: how different/better/relates to other work
+% Paragraph 4: What are the differences between your work, and what others
+% have done? Keep this at a high level, as you can refer to future sections
+% where specific details and differences will be given, but it is important
+% for the reader to know what is new about this work compared to other work
+% in the area.
 
-% para. 5: "We structure the remainder of this paper as follows."
+Our contributions are ...
+A discussion of the needs of real-time applications, and the appropriate
+transport services and APIs to support those needs.
+An outline of how those transport services can be realised over in the
+context of UDP and TCP.
+Preliminary deployments results, to show that there is the possibility
+that these services are usable in the public Internet.
+
+% Paragraph 5: "We structure the remainder of this paper as follows." Give
+% the reader a road-map for the rest of the paper. Try to avoid redundant
+% phrasing, "In Section 2, In section 3, ..., In Section 4, ... ", etc.
 
@@
 %==================================================================================================
-% Concentrate single-mindedly on a narrative that:
-%  - Describes the problem, and why it's interesting
-%  - Describes your idea
-%  - Defends your idea, showing how it solves the problem, and filling out
-%    the details
-% On the way, cite relevant work in passing, but defer discussion to the
-% end.
-%
-% Introduce the problem, and your idea, using examples, and only then
-% present the general case. Explain the idea as if your were speaking to
-% someone using a whiteboard. Conveying the intuition is primary; details
-% follow. Write in a top down manner: state broad themes and ideas first,
-% then go into details.
-%
-% The introduction makes claims. The body of the paper provides evidence
-% to support each claim. Check each claim in the introduction, identify
-% the evidence, and forward-reference it from the claim. 
-%==================================================================================================
-\section{}
+\section{Real-time Transport Services}
 
-% The problem
+...discussion of APIs and transport services for real-time applications
+
+- deadlines
+- partial reliability
+- application-level framing and ADUs
+- sub-streams
+- etc.
+
+what a conceptual API for this would look like
 
 %==================================================================================================
-\section{}
+\section{Innovation and Ossification}
 
-% My idea
+Briefly discuss: the Internet architecture, in principle, allows free
+innovation at the transport layer, provided the IP layer is unchanged. 
+Outline why this is not true in practice.
+
+Implication: new protocols must look like TCP or UDP.
+
+UDP is the obvious choice, since it provides minimal additional services
+over the IP layer, allowing great flexibility in innovation for protocols
+tunnelled on top. Only real cost is a few bytes of additional header.
+
+But, UDP is often blocked. Corporate firewalls. History of abuse for DDoS
+attacks (draft-byrne-opsec-udp-advisory-00). Slowly changing, e.g., due to
+QUIC, WebRTC, but not clear that UDP is universally available.
+
+TCP is a more complex choice for innovation. More sophisticated protocol.
+Complex headers. Mandates much more behaviour.
+
+However:
+- end-points can change congestion control, provided it's ACK-based
+- segments need be sent with contiguous sequence number space, resending
+  segments to fill gaps in sequence number space caused by packet loss
+- traditional API is restrictive, but flexibility to do out-of-order 
+  delivery to applications
+- easy to perform consistent segmentation on sending, and add COBS framing, 
+  to get message-oriented semantics
+
+Main issue with adding what we need is reliability. 
+Leads to the question: can we work-around TCP reliability?
 
 %==================================================================================================
-\section{}
+\section{Partial Reliability and TCP}
 
-% The details
+approach: 
+- separate out sequence numbers in sent segments with sequence numbers
+  visible to the application.
+- COBS encode for message boundaries.
+- retransmission have same segment sequence number, but different contents
+  and application sequence numbers
 
-% Describe results carefully:
-%  - clearly state assumptions
-%  - give enough information to allow the reader to recreate the results
-%  - ensure results are representative; statistically meaningful, etc.
-%  - don't overstate results
-%  - equally, don't understate them: consider the broader implications
+Does it work?
+- Initial deployment results from the UK
+- Is there any more data that can be got quickly?
+
+Any other surveys showing the same thing?
+
+%==================================================================================================
+\section{Realising Transport Services}
+
+Assuming the above is true more generally, once wider measurements have
+been conducted...
+
+here is an abstract API for real-time transport services
+
+we can clearly layer it above UDP: the services look a lot like PR-SCTP,
+and experiences with the WebRTC data channel show that this can run over
+UDP and be deployed in the wild.
+
+we also assert that these services can be deployed over TCP. How? TCP
+Hollywood offers a prototype. 
+brief summary.
+obviously we need to change TCP congestion control too.
 
@@
 
+to make effective use of the network, need to deploy new transport services
+and protocols
+
+seems likely that the appropriate long-term approach for doing this is to
+repurpose UDP as a demultiplexing layer for higher-layer protocols.
+
+in the short term, we need to deal with networks that block UDP, and be
+able to run over TCP.
+our initial work shows that this is frequently possible, and works well
+enough to offer benefit.
+
 %==================================================================================================

### 2016-05-10 paper.tex
Update re Marwan's comment

@@
 
-we also assert that these services can be deployed over TCP. How? TCP
-Hollywood offers a prototype. 
-brief summary.
-obviously we need to change TCP congestion control too.
+these services are deployed over TCP now, but have an inconvenient API that
+imposes lots of work on application developers, and with higher than desired
+latency. 
+TCP Hollywood offers a way to fix both issues.
+Brief summary.
+Obviously we need to change TCP congestion control too.
 

### 2016-05-10 paper.tex
Update pdfinfo

@@
 \pdfinfo{
-  /Title        (...)
-  /Author       (...)
-  /Subject      (...)
-  /Keywords     (..., ..., ...)
-  /CreationDate (D:20150827110616+01'00')
-  /ModDate      (D:20150827110616+01'00')
+  /Title        (Implementing Real-time Transport Services over an Ossified Network)
+  /Author       (Stephen McQuistin, Colin Perkins, and Marwan Fayed)
+  /Subject      (Real-time Transport Protocols)
+  /Keywords     (Transport Services, TCP, TCP Hollywood)
+  /CreationDate (D:20160516000000+01'00')
+  /ModDate      (D:20160516000000+01'00')
   /Creator      (LaTeX)

### 2016-05-15 paper.tex
First complete draft of the Introduction

@@
 ordered, byte stream API. These are suitable for many applications, but are
-not necessarily appropriate for real-time traffic.  UDP exposes essentially
-the raw IP packet delivery service, offering flexibility to develop new
-protocols, but at the cost of requiring completely new protocol mechanisms
-to be defined. While both protocols are usable for real-time applications,
-neither really provides the right services and API.
+not necessarily appropriate for real-time traffic.  UDP exposes the best-effort 
+IP packet delivery service, offering flexibility to develop new protocols,
+but at the cost of requiring completely new protocol mechanisms to be
+defined. Both protocols are usable for real-time applications, but neither
+really provides the right services and API.  This forces each application
+to re-invent mechanisms that should be provided by the transport, increasing 
+costs and complexity, raising barriers to innovation.
 
@@
 
-In this paper we develop an appropriate set of transport services and APIs
+In this paper we discuss the appropriate set of transport services and APIs
 for real-time applications.
-Then, we demonstrate that it is possible to realise these services and APIs
+We demonstrate that it is possible to realise these services and APIs
 in the context of both TCP and UDP, despite the limitations imposed by the
@@
 
-Our contributions are ...
-A discussion of the needs of real-time applications, and the appropriate
-transport services and APIs to support those needs.
-An outline of how those transport services can be realised over in the
-context of UDP and TCP.
-Preliminary deployments results, to show that there is the possibility
-that these services are usable in the public Internet.
+Our contributions are to identify the needs of real-time applications, and
+the appropriate transport services and APIs to support those needs; 
+to illustrate how those transport services can be realised on the current
+Internet, in the context of UDP and TCP deployments; and 
+to present some initial measurement results to show that the mechanisms we
+propose ought to be usable in the public Internet.
 
@@
 
-We structure the remainder of this paper as follows.
+We begin, in Section \ref{sec:services} by discussing transport services 
+for real-time applications, and outlining the common conceptual API that
+those applications use. 
+This is followed, in Section \ref{sec:ossification}, by a review of the
+deployment considerations for new protocols, caused by ossification of
+the network, with Section \ref{sec:partial} considering, in particular, how
+TCP reliability semantics can evolve within the constraints of the deployed
+infrastructure.
+Section \ref{sec:realising} outlines how the transport services we have
+identified can be realised in practical networks.
+Finally, Section \ref{sec:related} discusses related work, and Section
+\ref{sec:conclusions} concludes.
 
@@
 \section{Real-time Transport Services}
+\label{sec:services}
 
@@
 \section{Innovation and Ossification}
+\label{sec:ossification}
 
@@
 \section{Partial Reliability and TCP}
+\label{sec:partial}
 
@@
 \section{Realising Transport Services}
+\label{sec:realising}
 
@@
 \section{Related Work}
+\label{sec:related}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2016-05-15 paper.tex
Expand discussion of real-time transport services

@@
 
-...discussion of APIs and transport services for real-time applications
+In the IETF, the Transport Services (TAPS) working group is chartered to 
+(1) develop a taxonomy of \emph{transport services}, i.e., to identify the
+features that comprise, and can be combined, to form complete transport
+protocols; and (2) to develop an abstract API for applications to request
+desirable services, allowing the system to select an appropriate transport
+protocol based on application needs. It is hoped that this will loosen the
+coupling between application and transport, so enabling deployment of new 
+transport protocols.
 
-- deadlines
-- partial reliability
-- application-level framing and ADUs
-- sub-streams
-- etc.
+...TAPS is valuable, because it gives us a vocabulary for discussing the
+   components of transport protocols, that we can use to identify the needs
+   of real-time applications
 
-what a conceptual API for this would look like
+...the key feature of such applications is clearly timing. That is, since
+   these applications convey real-time data, they all have some concept 
+   of a \emph{deadline} by which that data must be presented, after which it 
+   becomes useless. 
+   the deadline might be short or long. Interactive applications, such as
+   telephony, video conferencing, or telepresence, require low end-to-end
+   latency, with the deadline for presenting data being tens, or perhaps
+   low hundreds, of milliseconds after it was generated. Non-interactive
+   applications, such as broadcast TV, may have deadlines in the order of
+   seconds, or even tens of seconds for on-demand programming. 
+   These deadlines are unusual for real-time systems, in that they are
+   simultaneously flexible and strict: flexible in that the exact value
+   of the deadline is typically not important, provided it is of the 
+   right order-of-magnitude for the application, but strict in that 
+   any particular deadline provides a hard cut-off, after which the
+   data is useless.
+
+...the presence of deadlines, in a network that offers only a best-effort
+   packet deliver service, implies that the transport must provide only
+   \emph{partial reliability}.
+   the network must be assumed to have non-zero probability of losing any
+   particular packet. If FEC is used to repair loss, this implies there is
+   always some probability that a particular packet will be non-recoverable. 
+   If retransmission is used to repair loss, this implies that there is
+   always the potential of unbounded delay. 
+   These probabilities can be estimated and bounded, but will be non-zero.
+   Accordingly, to meet deadlines, the transport must offer a partial
+   reliability service. 
+   Many real-time applications run over TCP today, and TCP does not offer
+   this service. This results in occasional play-out stalls, when the
+   transport blocks the application. These are one of the primary causes
+   of poor user experience in streaming applications, and this is directly
+   caused by the lack of partial reliability. Missing one frame that has
+   not been delivered by its deadline is much less disruptive than a stall
+   in play-out.
+
+...given deadlines and partial reliability, the next important transport
+   service is \emph{dependency management}. There is no point sending
+   data that cannot be used, because it depends on previous data that
+   was lost. 
+   complex trade-off, since you might want to send data that misses its 
+   deadline, if it allows later data to be used.
+
+...following from this, we have the concept of application-level framing
+   to make the best use of data, irrespective of losses. This implies that
+   the transport provides a \emph{message oriented} service that respects
+   ADU boundaries.
+
+...we note that many multimedia applications make use of multiple streams. 
+   For example, audio and video are logically separate, etc.
+   can be supported by opening multiple transport connections, but
+   desirable to support a \emph{sub-stream} service to reduce overheads.
+
+...finally, we note that \emph{congestion control} is important. 
+   motivate: historically there has been a belief that real-time doesn't 
+   do congestion control, needs an isochronous channel. 
+   not practical for the Internet. 
+   Also, experience has shown not necessary for many real-time applications.
+   Yes, there are some that need constant bandwidth and cannot adapt, but
+   many are adaptive. 
+   Better user experience to adapt to fit the channel, rather than recover
+   from loss after the channel discards packets that don't fit.
+
+..what a conceptual API for this would look like?
 

### 2016-05-16 paper.tex
Revise Real-time Transport Services

--- 2016-05-15 paper.tex
@@
 
-...discussion of APIs and transport services for real-time applications
+In the IETF, the Transport Services (TAPS) working group is chartered to 
+(1) develop a taxonomy of \emph{transport services}. That is, to identify the
+features that comprise, and can be combined, to form complete transport
+protocols; and (2) to develop an abstract API for applications to request
+desirable services, allowing the system to select an appropriate transport
+protocol based on application needs. It is hoped that this will loosen the
+coupling between application and transport, so enabling deployment of new 
+transport protocols.
 
-- deadlines
-- partial reliability
-- application-level framing and ADUs
-- sub-streams
-- etc.
+The work in TAPS is valuable, because it gives us a vocabulary for
+discussing the components of transport protocols.  We can then use that
+vocabulary when discussing the needs of real-time applications, and the
+protocols needed to support those needs.
 
-what a conceptual API for this would look like
+The essential feature of real-time applications is clearly timing. That is,
+since these applications convey real-time data, they all have some concept
+of a \emph{deadline} by which that data must be presented, after which it
+becomes useless. 
+The deadline might be short or long. Interactive applications, such as
+telephony, video conferencing, or telepresence, require low end-to-end
+latency, with the deadline for presenting the media (playing the audio,
+displaying the video frame) being tens, or perhaps low hundreds, of
+milliseconds after it was captured. Non-interactive applications, such as
+broadcast TV, may have deadlines in the order of seconds, or even tens of
+seconds for on-demand programming. 
+The deadlines in many networked multimedia applications are unusual when
+compared to other real-time systems. They are simultaneously flexible and
+strict: flexible in that the exact value of the deadline is typically not
+important, provided it is of the right order-of-magnitude for the
+application, but strict in that any particular deadline provides a hard
+cut-off, after which the data is useless.
+
+The presence of deadlines, in a network that offers only a best-effort
+packet deliver service, implies that the transport must provide
+\emph{partial reliability}.
+The network has non-zero probability of losing any particular packet. 
+If forward error correction is used to repair loss, this implies there is
+always some probability that a particular packet will be non-recoverable.
+If retransmission is used to recover from losses, there is the potential 
+for unbounded delay, since any retransmission may itself be lost. 
+These probabilities can be estimated, to give probabilistic constraints
+on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
+the transport must support partial reliability service, accepting that
+it may not be able to deliver all data by its deadline.
+Many real-time applications run over TCP today, and TCP does not offer a
+partial reliability service. This results in occasional play-out stalls, 
+when the transport blocks the application because a retransmission takes
+too long. These stalls are one of the primary causes of poor user
+experience in streaming applications, and are directly caused by the lack
+of partial reliability in TCP. Missing one frame that has not been
+delivered by its deadline is much less disruptive than a stall in play-out.
+
+Given deadlines and partial reliability, the next important transport
+service is \emph{dependency management}. There is no point sending data
+that cannot be used, because it depends on previous data that was lost. 
+...complex trade-off, since you might want to send data that misses its 
+   deadline, if it allows later data to be used.
+
+...following from this, we have the concept of application-level framing
+   to make the best use of data, irrespective of losses. This implies that
+   the transport provides a \emph{message oriented} service that respects
+   ADU boundaries.
+
+...we note that many multimedia applications make use of multiple streams. 
+   For example, audio and video are logically separate, etc.
+   can be supported by opening multiple transport connections, but
+   desirable to support a \emph{sub-stream} service to reduce overheads.
+
+...finally, we note that \emph{congestion control} is important. 
+   motivate: historically there has been a belief that real-time doesn't 
+   do congestion control, needs an isochronous channel. 
+   not practical for the Internet. 
+   Also, experience has shown not necessary for many real-time applications.
+   Yes, there are some that need constant bandwidth and cannot adapt, but
+   many are adaptive. 
+   Better user experience to adapt to fit the channel, rather than recover
+   from loss after the channel discards packets that don't fit.
+
+..what a conceptual API for this would look like?
 

### 2016-05-16 paper.tex
Expand Real-time Transport Services

@@
 
-\newcommand{\todo}[1]{\textbf{\textcolor{red}{To do: #1}}}
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{To do -- #1}}}
 
@@
 
-..what a conceptual API for this would look like?
+We note that a connection-oriented transport is not a core requirement for
+many real-time multimedia applications.  Indeed, flexibility to change the
+destination of data within a call could be beneficial for applications that
+support mobile users, or for some forms of multiparty session. However, to
+support NAT traversal and to help dynamically manage firewall pinholes, it
+is often desirable for the transport to be connection oriented. We find
+these concerns outweigh the benefits of connectionless transport, and so
+add a requirement for connection oriented service. 
+Similarly, while not strictly needed by the applications, it is beneficial
+if the transport provides a keep-alive service to refresh NAT and firewall
+bindings if the application goes silent.
+
+Table \ref{tab:services} summarises the transport services required for
+real-time multimedia applications.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ll}
+    \toprule
+      Transport Service     & Requirement \\
+    \midrule
+      Deadlines             & Core        \\
+      Partial reliability   & Core        \\
+      Dependencies          & Core        \\
+      Message-oriented      & Core        \\
+      Sub-streams           & Core        \\
+      Congestion controlled & Core        \\
+      Connection oriented   & Subsidiary  \\
+      Keep-alive            & Subsidiary  \\
+    \bottomrule
+  \end{tabular}
+  \caption{Transport services required for real-time multimedia}
+  \label{tab:services}
+\end{table}
+
+Given a transport protocol that provides the above services, how could the
+API between transport and application be structured?
+
+\todo{Stephen: outline the TCP Hollywood API, with text and a figure}
+
+It is to be noted that this API is not dissimilar to the PR-SCTP abstract
+API. \todo{CSP: check this}
 

### 2016-05-16 paper.tex
Work on Innovation and Ossification

@@
 
+\begin{figure}
+  \centering
+  ...
+  \caption{Outline transport API for real-time applications}
+  \label{fig:api}
+\end{figure}
+
 It is to be noted that this API is not dissimilar to the PR-SCTP abstract
@@
 
-Briefly discuss: the Internet architecture, in principle, allows free
-innovation at the transport layer, provided the IP layer is unchanged. 
-Outline why this is not true in practice.
+The Internet architecture, in principle, allows free innovation at the
+transport layer, provided the underlying network (IP) layer is unchanged.  
+Routers should inspect the source addresses of packets to perform network
+ingress filtering \cite{RFC2827}, and the destination addresses to route
+packets to the correction destination, but should not inspect their
+contents. This is not, of course, how the real network operates.
 
-Implication: new protocols must look like TCP or UDP.
+\todo{outline why this is not true in practice}
 
-UDP is the obvious choice, since it provides minimal additional services
-over the IP layer, allowing great flexibility in innovation for protocols
-tunnelled on top. Only real cost is a few bytes of additional header.
+The implication of this reality is that it is difficult to deploy new 
+transport protocols. The installed base of NATs, firewalls, and other
+middleboxes is such that packets that do not look like TCP or UDP are
+unlikely to pass the network. We may innovate all we like, provided the
+transport of the future looks like TCP or UDP to middleboxes.\footnote{
+  This is inconvenient, certainly, but is not necessarily a bad thing. 
+  The Internet is critical infrastructure. It support emergency services, 
+  healthcare applications, infrastructure components, financial services, 
+  and so on, many of which are essential to the functioning of society. 
+  Making changes to this type of infrastructure \emph{should} require 
+  careful backwards compatibility.
+}
 
-But, UDP is often blocked. Corporate firewalls. History of abuse for DDoS
-attacks (draft-byrne-opsec-udp-advisory-00). Slowly changing, e.g., due to
-QUIC, WebRTC, but not clear that UDP is universally available.
+UDP is the obvious base for future protocol development, since it provides
+minimal additional services over the IP layer, allowing great flexibility
+in innovation for protocols tunnelled on top. Provided middleboxes do not
+inspect the payload too carefully, the only real cost to innovation, when
+compared to a native transport protocol running over IP, is a few bytes of
+additional header. Examples in this space include RTP \cite{RFC3550}, one
+of the most widely deployed real-time transport protocols; the WebRTC Data
+Channel \cite{}, which tunnels peer-to-peer SCTP associations over a DTLS
+association over UDP; and QUIC \cite{}, which provides a modern alternative 
+to TCP, implemented over UDP.
 
-TCP is a more complex choice for innovation. More sophisticated protocol.
-Complex headers. Mandates much more behaviour.
+Despite these advantages, UDP can be problematic as a substrate for new
+protocol development. UDP traffic is blocked by some enterprise firewalls,
+and some in the operations community have a strong distrust of UDP-based
+protocols and applications \cite{draft-byrne-opsec-udp-advisory-00}. In
+part this is due to ignorance. Outside specific niches, such as DNS, UDP
+has not been widely used in enterprise environments, and hence is widely
+misunderstood. Blocking the unknown is a rational response. In addition,
+UDP traffic has been widely used as a component of distributed denial of
+service (DDoS) attacks, leading some to install blanket blocks of UDP as 
+a safety measure (blanket blocking, rather than the more targeted blocks 
+used when TCP traffic is used in DDoS attacks, are justified using the
+argument that UDP is not widely used).
+These issues are slowly changing, as UDP-based applications penetrate the
+enterprise consciousness, but not clear that UDP is universally available
+(Google report XX\% \todo{how much?} of endpoints are reachable with QUIC
+running over UDP \cite{}).
 
-However:
-- end-points can change congestion control, provided it's ACK-based
+TCP is a more complex choice for innovation. It is a more sophisticated 
+protocol than UDP, with complex headers, and a protocol state machine
+that mandates much more behaviour and is widely understood, and policed,
+by in-network middleboxes. This does not mean that TCP cannot evolve, or
+form the basis for new transport services. Rather, it means that any
+innovation or development must be done carefully, paying very careful
+attention to backwards compatibility.
+
+We identify a number of places where TCP can evolve with comparative
+freedom. 
+Firstly, the TCP congestion control algorithm is executed by the end
+points, and can be changed provided the new version requires no new
+information to be exchanged. 
+\todo{expand, and discuss how this evolution has occurred to date}
+
+Then, we note that the API that is exposed to applications using TCP is
+not visible to the network. 
+- traditional API is restrictive, but flexibility to do out-of-order 
+  delivery to applications
+
+- easy to perform consistent segmentation on sending, and add COBS framing, 
+  to get message-oriented semantics
+
 - segments need be sent with contiguous sequence number space, resending
   segments to fill gaps in sequence number space caused by packet loss
-- traditional API is restrictive, but flexibility to do out-of-order 
-  delivery to applications
-- easy to perform consistent segmentation on sending, and add COBS framing, 
-  to get message-oriented semantics
 

### 2016-05-16 paper.tex
Finish sketching out Innovation and Ossification

@@
 We identify a number of places where TCP can evolve with comparative
-freedom. 
-Firstly, the TCP congestion control algorithm is executed by the end
-points, and can be changed provided the new version requires no new
-information to be exchanged. 
+freedom. These include congestion control, the end-point API, and data
+segmentation. If care is taken, there is also the possibility to change
+the reliability semantic.
+
+The TCP congestion control algorithm is executed by the end points, and 
+can be changed, provided the new version requires no new information to 
+be exchanged. 
 \todo{expand, and discuss how this evolution has occurred to date}
+We note that, while standardised TCP congestion control has followed the
+goal of maximising throughput at the expense of latency and variability,
+this is not required by the protocol. TCP Vegas \cite{brakmo:1994:tcp-vegas}
+is perhaps the best known approach that changes these constraints, with a
+delay-based algorithm that reduces latency. It would also be possible to
+implement alternatives that seek stability, or compatibility with the
+dictates of a video codec, rather than the traditional ``TCP Friendly''
+congestion control -- even if implemented within TCP.
 
-Then, we note that the API that is exposed to applications using TCP is
-not visible to the network. 
-- traditional API is restrictive, but flexibility to do out-of-order 
-  delivery to applications
+The API that is exposed to applications using TCP is not visible to the
+network, and can be changed. 
+\todo{discuss how the API can be relaxed to allow out-of-order delivery}
+\todo{discuss how the API can be modified to accept/delivery a sequence
+      of messages, rather than a byte stream, and how with consistent
+      segmentation at the sender and COBS framing, this can be efficient}
 
-- easy to perform consistent segmentation on sending, and add COBS framing, 
-  to get message-oriented semantics
-
-- segments need be sent with contiguous sequence number space, resending
-  segments to fill gaps in sequence number space caused by packet loss
-
-Main issue with adding what we need is reliability. 
-Leads to the question: can we work-around TCP reliability?
+These changes could address many of the transport service needs we have for
+real-time applications, but still leave a critical issue of how to improve
+timing behaviour. Specifically, how to enable partial reliability for TCP,
+after which it is possible to layer-on support for managing deadlines and 
+dependencies.
 

### 2016-05-17 paper.tex
Adds QUIC connectivity number

@@
 It is to be noted that this API is not dissimilar to the PR-SCTP abstract
-API. \todo{CSP: check this}
+API, which provides \textit{timed reliability}, using a ``lifetime''
+specified by the application. \todo{CSP: check this}
 
@@
 enterprise consciousness, but not clear that UDP is universally available
-(Google report XX\% \todo{how much?} of endpoints are reachable with QUIC
-running over UDP \cite{}).
+(Google report 90-95\% of endpoints are reachable with QUIC
+running over UDP \cite{roskind2013quic}).
 

### 2016-05-17 paper.tex
Add comment about QUIC results

@@
 (Google report 90-95\% of endpoints are reachable with QUIC
-running over UDP \cite{roskind2013quic}).
+running over UDP \cite{roskind2013quic}, but it is not clear that the set
+of hosts running their Chrome browser is representative of all Internet
+environments).
 

### 2016-05-18 paper.tex
Drafts 'Partial Reliability and TCP' section

--- 2016-05-17 paper.tex
@@
   Making changes to this type of infrastructure \emph{should} require 
-  careful backwards compatibility.
+  careful backwards compatibility \cite{mcquistin:2015:reinterpreting}.
 }
@@
 
-approach: 
-- separate out sequence numbers in sent segments with sequence numbers
-  visible to the application.
-- COBS encode for message boundaries.
-- retransmission have same segment sequence number, but different contents
-  and application sequence numbers
+Partial reliability (i.e., reliability conditional on timing and dependency
+information) can be implemented by relaxing TCP's reliability guarantee.
+The implication of this is that we need to offer a message-oriented
+abstraction to applications. If the arrival of a segment cannot be
+guaranteed, then it is not possible to offer a byte stream abstraction.
 
-Does it work?
-- Initial deployment results from the UK
-- Is there any more data that can be got quickly?
+To offer a message-oriented abstraction, the boundaries between each
+message must be maintained between sender and receiver. This means that a
+framing mechanism is required: it is not sufficient to send each message in
+a single segment, as this mapping will not necessarily be maintained by the
+network. A framing marker is added to the start and end of each message
+before transmission. An encoding algorithm is used to escape all occurences
+of the framing marker within the message data. This process does not impact
+on the data that can be sent or received by applications.
 
-Any other surveys showing the same thing?
+In order to maintain compatibility with middleboxes, offering partial
+reliability requires using \emph{inconsistent retransmissions}. This means
+that the mapping between message data and TCP sequence numbers is no longer
+static: a given TCP sequence number may be relate to different messages at
+different times. Therefore, an application-level sequence number is
+required to allow messages to be uniquely identified.
+
+When a TCP segment is to be retransmitted, the current mapping between its
+sequence number and application-level sequence numbers is used to determine
+which messages within the segment are to be retransmitted. A liveness check
+is performed on these messages, to determine that (i) the message will
+arrive on time to be played out; and (ii) the message does not depend on an
+expired message. For (i), we combine the time that the message has spent in
+a sending queue, with an estimate of the round-trip time and the current
+play-out delay. This is then compared against the lifetime of the message,
+as expressed by the application. For (ii), we maintain metadata about
+sequence numbers that have expired, and check this metadata for the
+dependency expressed by the application.
+
+This mechanism -- inconsistent retransmissions -- is visible to middleboxes
+on the network that are performing payload inspection. These middleboxes
+may interpret this behaviour relating to an attack. For example, a
+man-on-the-side attack exhibits similar behaviour, where a malicious host
+is injecting data into an existing TCP flow. As a result, our connection
+may be disrupted. Honda et al. \cite{honda:2011:extend-tcp} conducted
+experiments across 135 paths on the Internet, to determine support for
+inconsistent retransmissions. The observed that the majority of paths
+delivered inconsistent retransmissions successfully. On Port 80 (HTTP), the
+original segment was delivered on 7\% of paths tested. Only one connection
+reset was observed. We conducted further deployability experiments using
+inconsistent retransmissions \cite{mcquistin2016hollywood}, and found that
+fixed-line networks delivered inconsistent retransmissions successfully,
+but that delivery of the original segment was common on cellular networks.
+
+These deployability experiments suggest that our protocol should be
+flexible: inconsistent retransmissions may not be delivered, and we should
+handle reception of the original segment. If the protocol detects that
+inconsistent retransmissions are not being delivered, then the mechanism
+can be disabled for the connection. Further, if a connection reset occurs,
+then the connection should be retried with the mechanism disabled.
+
+\todo{Add discussion on checksumming to maintain message integrity}
 

### 2016-05-21 paper.tex
Adds outline, abstract API and notes in other sections

@@
 Table \ref{tab:services} summarises the transport services required for
-real-time multimedia applications.
+real-time multimedia applications. From this set of services, Table
+\ref{tab:api} gives an outline API. As shown, the connection primitives
+are the same as those of TCP. Hosts setup and teardown sockets using 
+\texttt{create\_socket()} and \texttt{close\_socket()}. Servers
+\texttt{bind()} to an address they wish to accept connections from, they
+\texttt{listen()} on for incoming connections, and they \texttt{accept()}
+an incoming connection. Clients \texttt{connect()} to a server.
+
+The message-oriented service is exposed by the \texttt{send\_message()}
+and \texttt{receive\_message()} functions. When sending messages,
+applications specify a sequence number. Applications
+set the receiver's play-out delay, using \texttt{set\_po\_delay()}, in
+milliseconds. This is sent to the sender, for use as part of the deadline
+estimation. The sender attaches a \texttt{lifetime} to each message,
+which is combined with an estimate of the current round-trip-time, and the
+time that the message has spent in the sending buffer, to determine if a
+message will arrive within its lifetime. A message that won't arrive
+within its lifetime is considered to have \emph{expired}. Applications
+specify the application-level sequence number of a message they depend on
+(\texttt{depends\_on}). If the message depends on a message that has
+expired, then it too is considered to have expired. The partial
+reliability service follows from the deadline and dependency services:
+messages will be reliably transmitted until they expire.
 
@@
 
-Given a transport protocol that provides the above services, how could the
-API between transport and application be structured?
-
-\todo{Stephen: outline the TCP Hollywood API, with text and a figure}
-
-\begin{figure}
+\begin{table}
   \centering
-  ...
+  \begin{tabular}{ll}
+    \toprule
+      Function & Parameters \\
+    \midrule
+      \texttt{create\_socket} &  \\
+      \texttt{close\_socket} & \texttt{sd} -- Socket descriptor \\
+    \midrule
+      \texttt{bind} & \texttt{sd} -- Socket descriptor \\
+                    & \texttt{addr} -- Address to bind to \\
+                    & \texttt{addrlen} -- Length of \texttt{addr} \\
+      \texttt{listen} & \texttt{sd} -- Socket descriptor \\
+            \texttt{accept} & \texttt{sd} -- Socket descriptor \\
+                    & \texttt{addr} -- Address of peer \\
+                    & \texttt{addrlen} -- Length of \texttt{addr} \\
+      \texttt{connect} & \texttt{addr} -- Address to connect to \\
+                    & \texttt{addrlen} -- Length of \texttt{addr} \\
+    \midrule
+      \texttt{set\_po\_delay} & \texttt{delay} -- Playout delay (in ms) \\
+    \midrule
+      \texttt{send\_message} & \texttt{sd} -- Socket descriptor \\
+      & \texttt{buf} -- Message data \\
+      & \texttt{len} -- Length of message data \\
+      & \texttt{seq\_num} -- Sequence number \\
+      & \texttt{lifetime} -- Lifetime of message (in ms) \\ 
+      & \texttt{depends\_on} -- \texttt{seq\_num} of dependency \\
+      & \texttt{substream} -- Substream identifier \\
+      \texttt{receive\_message} & \texttt{sd} -- Socket descriptor \\
+      & \texttt{buf} -- Buffer for message data \\
+      & \texttt{len} -- Size of \texttt{buf} \\
+      & \texttt{substream} -- Substream identifier \\
+    \bottomrule
+  \end{tabular}
   \caption{Outline transport API for real-time applications}
-  \label{fig:api}
-\end{figure}
+  \label{tab:api}
+\end{table}
 
@@
 
+\todo{Other motivation for using TCP: HTTP-based streaming protocols}
+
 TCP is a more complex choice for innovation. It is a more sophisticated 
@@
 The API that is exposed to applications using TCP is not visible to the
-network, and can be changed. 
-\todo{discuss how the API can be relaxed to allow out-of-order delivery}
+network, and can be changed. Relaxing the API to enable the out-of-order
+delivery of segments is trivial: segments are delivered to the application
+in the order that they arrive, with their TCP sequence attached. The TCP
+sequence number can be passed to the application using the existing
+Berkeley sockets API, either with the received data, or using 
+\texttt{getsockopt()}. Out-of-order delivery is not useful when using a 
+byte-stream abstraction, and so the API should be further modified to
+provide a message-oriented abstraction. The Berkeley sockets API already
+supports such an abstraction for datagram protocols.
+
 \todo{discuss how the API can be modified to accept/delivery a sequence
@@
 
-Assuming the above is true more generally, once wider measurements have
-been conducted...
+While further measurement studies are required to determine the
+deployability of wire-visible changes to TCP (such as inconsistent
+retransmissions), we have shown that we can provide all of the transport 
+services needed by real-time applications, using either TCP or UDP. The
+API shown in Table \ref{tab:api} is largely similar to that PR-SCTP, which
+has seen real-world deployments within the WebRTC protocol, demonstrating
+that we can layer these services above UDP.
 
-here is an abstract API for real-time transport services
-
-we can clearly layer it above UDP: the services look a lot like PR-SCTP,
-and experiences with the WebRTC data channel show that this can run over
-UDP and be deployed in the wild.
+However, as described in Section \ref{sec:ossification}, there are
+deployability advantages to providing these services across TCP.
 

### 2016-05-22 paper.tex
Expand API discussion

@@
 In the IETF, the Transport Services (TAPS) working group is chartered to 
-(1) develop a taxonomy of \emph{transport services}. That is, to identify the
+(1) develop a taxonomy of \emph{transport services}, that is, to identify the
 features that comprise, and can be combined, to form complete transport
@@
 transport protocols.
+
+%--------------------------------------------------------------------------------------------------
+\subsection{Transport Services}
 
@@
 that cannot be used, because it depends on previous data that was lost. 
-...complex trade-off, since you might want to send data that misses its 
-   deadline, if it allows later data to be used.
+\todo{...complex trade-off, since you might want to send data that misses its 
+   deadline, if it allows later data to be used.}
 
-...following from this, we have the concept of application-level framing
+\todo{...following from this, we have the concept of application-level framing
    to make the best use of data, irrespective of losses. This implies that
    the transport provides a \emph{message oriented} service that respects
-   ADU boundaries.
+   ADU boundaries.}
 
-...we note that many multimedia applications make use of multiple streams. 
+\todo{...we note that many multimedia applications make use of multiple streams. 
    For example, audio and video are logically separate, etc.
    can be supported by opening multiple transport connections, but
-   desirable to support a \emph{sub-stream} service to reduce overheads.
+   desirable to support a \emph{sub-stream} service to reduce overheads.}
 
-...finally, we note that \emph{congestion control} is important. 
+\todo{...finally, we note that \emph{congestion control} is important. 
    motivate: historically there has been a belief that real-time doesn't 
@@
    Better user experience to adapt to fit the channel, rather than recover
-   from loss after the channel discards packets that don't fit.
+   from loss after the channel discards packets that don't fit. }
 
@@
 Table \ref{tab:services} summarises the transport services required for
-real-time multimedia applications. From this set of services, Table
-\ref{tab:api} gives an outline API. As shown, the connection primitives
-are the same as those of TCP. Hosts setup and teardown sockets using 
-\texttt{create\_socket()} and \texttt{close\_socket()}. Servers
-\texttt{bind()} to an address they wish to accept connections from, they
-\texttt{listen()} on for incoming connections, and they \texttt{accept()}
-an incoming connection. Clients \texttt{connect()} to a server.
+real-time multimedia applications.
 
-The message-oriented service is exposed by the \texttt{send\_message()}
-and \texttt{receive\_message()} functions. When sending messages,
-applications specify a sequence number. Applications
-set the receiver's play-out delay, using \texttt{set\_po\_delay()}, in
-milliseconds. This is sent to the sender, for use as part of the deadline
-estimation. The sender attaches a \texttt{lifetime} to each message,
-which is combined with an estimate of the current round-trip-time, and the
-time that the message has spent in the sending buffer, to determine if a
-message will arrive within its lifetime. A message that won't arrive
-within its lifetime is considered to have \emph{expired}. Applications
-specify the application-level sequence number of a message they depend on
-(\texttt{depends\_on}). If the message depends on a message that has
-expired, then it too is considered to have expired. The partial
-reliability service follows from the deadline and dependency services:
-messages will be reliably transmitted until they expire.
+%--------------------------------------------------------------------------------------------------
+\subsection{Abstract API}
+
+Given the set of transport services outlined in Table \ref{tab:services}, 
+we sketch an abstract API in  Table \ref{tab:api}. The primitives divide
+into four categories:
+
+\begin{itemize}
+  \item Hosts setup and tear-down sockets using the \texttt{socket()}
+    and \texttt{close()} functions, as in the standard Berkeley sockets
+    API.
+
+  \item The connection primitives are the same as those of TCP sockets.
+    Servers \texttt{bind()} to a particular address and port, then
+    \texttt{listen()} for and \texttt{accept()} incoming connections.
+    Clients \texttt{connect()} to a server.
+
+   \item Once the connection is established, the receiver application
+     indicates the media play-out delay it is using, in milliseconds, via
+     the \texttt{set\_po\_delay()} call. This specifies the time that the
+     application will buffer data, to compensate for network timing jitter,
+     before it is rendered to the user. The play-out delay is fed back to
+     the sender host, for use as part of the media deadline estimation.
+
+  \item Finally, message-oriented data transmission is exposed by the
+    \texttt{send\_message()} and \texttt{recv\_message()} functions.  
+    These expose a partially reliable message delivery service to the
+    application, framing data such that either a complete message is
+    delivered, or it is lost in its entirety.
+
+    The \texttt{send\_message()} call takes four additional parameters
+    compared to the standard Berkeley Sockets API. These are 1) a message
+    sequence number, that can be used to re-order messages and detect 
+    message loss; 2) a relative deadline, which is combined with an
+    estimate of the current round-trip-time, and the time that the message
+    has spent in the sending buffer, to determine if a message will arrive
+    in time to be played-out; 3) the sequence number of any message on
+    which this depends, for example the sequence number of a video I-frame
+    on which a P-frame is predicted; and 4) a sub-stream identifier, used,
+    for example, to differentiate audio, video, sub-title, control, and
+    repair streams.
+
+    The \texttt{recv\_message()} call returns the sequence number,
+    dependency information, and sub-stream identifier along with any
+    received message, allowing the receiver to direct it to the correct
+    decoding queue.
+
+    \todo{relate this to the transport services described earlier}
+
+\end{itemize}
+
+A message that won't arrive within its lifetime is considered to have
+\emph{expired}. Applications specify the application-level sequence number
+of a message they depend on (\texttt{depends\_on}). If the message depends
+on a message that has expired, then it too is considered to have expired.
+The partial reliability service follows from the deadline and dependency
+services: messages will be reliably transmitted until they expire.
 
@@
     \midrule
-      \texttt{create\_socket} &  \\
-      \texttt{close\_socket} & \texttt{sd} -- Socket descriptor \\
+      \texttt{socket} & \texttt{af} -- Address family \\
+                      & \texttt{st} -- Socket type \\
+      \texttt{close} & \texttt{sd} -- Socket descriptor \\
     \midrule
@@
       & \texttt{seq\_num} -- Sequence number \\
-      & \texttt{lifetime} -- Lifetime of message (in ms) \\ 
+      & \texttt{deadline} -- Relative deadline of message (in ms) \\ 
       & \texttt{depends\_on} -- \texttt{seq\_num} of dependency \\
       & \texttt{substream} -- Substream identifier \\
-      \texttt{receive\_message} & \texttt{sd} -- Socket descriptor \\
+      \texttt{recv\_message} & \texttt{sd} -- Socket descriptor \\
       & \texttt{buf} -- Buffer for message data \\
@@
       & \texttt{substream} -- Substream identifier \\
+      & \todo{return value, rather than param?}\\
     \bottomrule
   \end{tabular}
-  \caption{Outline transport API for real-time applications}
+  \caption{Outline transport API for real-time applications 
+          \todo{show return values of these calls}}
   \label{tab:api}

### 2016-05-22 paper.tex
Note related work
Expand Realising Transport Services

@@
 network. A framing marker is added to the start and end of each message
-before transmission. An encoding algorithm is used to escape all occurences
+before transmission. An encoding algorithm is used to escape all occurrences
 of the framing marker within the message data. This process does not impact
@@
 original segment was delivered on 7\% of paths tested. Only one connection
-reset was observed. We conducted further deployability experiments using
+reset was observed.
+
+We conducted further deployment experiments using
 inconsistent retransmissions \cite{mcquistin2016hollywood}, and found that
@@
 but that delivery of the original segment was common on cellular networks.
+\todo{expand this}
 
-These deployability experiments suggest that our protocol should be
+These deployment experiments suggest that our protocol should be
 flexible: inconsistent retransmissions may not be delivered, and we should
@@
 
-While further measurement studies are required to determine the
-deployability of wire-visible changes to TCP (such as inconsistent
-retransmissions), we have shown that we can provide all of the transport 
-services needed by real-time applications, using either TCP or UDP. The
-API shown in Table \ref{tab:api} is largely similar to that PR-SCTP, which
-has seen real-world deployments within the WebRTC protocol, demonstrating
-that we can layer these services above UDP.
+While further measurement studies are required to confirm the ability to
+deploy wire-visible changes to TCP (such as inconsistent retransmissions),
+we have shown that we can provide all of the transport services needed by
+real-time applications, using either TCP or UDP.
 
-However, as described in Section \ref{sec:ossification}, there are
-deployability advantages to providing these services across TCP.
+Evidence that these services can be deployed above UDP exists in the form
+of the WebRTC data channel \cite{}.  
+\todo{expand}
 
+\todo{also QUIC}
+
+% The API shown in Table \ref{tab:api} is largely similar to that PR-SCTP,
+% which has seen real-world deployments within the WebRTC protocol,
+% demonstrating that we can layer these services above UDP.
+
+Deployments using UDP are popular, and work well. However, as described in
+Section \ref{sec:ossification}, there are also reasons for providing these 
+services running over TCP. 
+\todo{
   these services are deployed over TCP now, but have an inconvenient API that
@@
   Obviously we need to change TCP congestion control too.
+}
+
+Taken together, the wide experiences with the WebRTC Data Channel and QUIC
+demonstrate that the transport services necessary to support real-time
+traffic could be deployed running over UDP. Our work prototyping the TCP
+Hollywood protocol, and earlier measurements by Honda \emph{et al}.\
+\cite{} also suggest that deployment over TCP is possible.
+
+The challenge for the future is in combining such techniques below a common
+API, so that an application can transparently switch between UDP-based and
+TCP-based transport, depending on what is supported by the underlying
+network. This is the promise of the TAPS API, that we have shown ought to
+be feasible for real-time applications.
 
@@
 %  - Ensure related work is accurate and up-to-date
+
+\todo{Stephen: TCP minion; time-lined TCP}
+
+\todo{QUIC}
+
+\todo{SPUD}
+
+\todo{WebRTC data channel - SCTP over UDP}
+
+\todo{RFC6773 - DCCP over UDP}
 

### 2016-05-23 paper.tex
Adds TCP Minion and TLTCP to Related Work

@@
 
-\todo{Stephen: TCP minion; time-lined TCP}
+The Minion protocol suite \cite{nowlan:2012:minion} includes uTCP, a
+protocol that uses TCP as a substrate to provide an unordered,
+message-oriented service to applications. While this enables many of
+the transport services described in Section \ref{sec:services}, support for
+partial reliability, deadlines, and dependencies is not provided. 
+Time-Lined TCP (TLTCP) \cite{mukherjee:2000:timelines} similarly provides a
+message-oriented service, but allows applications to attach a time-line
+to messages. Messages are (re-)transmitted asunder standard TCP within
+their time-line, after which they are discarded. The mechanism by which
+this service is provided (i.e., the introduction of gaps in the sequence
+space) reduces deployability.
 

### 2016-05-23 paper.tex
Adds references

@@
 Routers should inspect the source addresses of packets to perform network
-ingress filtering \cite{RFC2827}, and the destination addresses to route
+ingress filtering \cite{rfc:2827}, and the destination addresses to route
 packets to the correction destination, but should not inspect their
@@
 compared to a native transport protocol running over IP, is a few bytes of
-additional header. Examples in this space include RTP \cite{RFC3550}, one
+additional header. Examples in this space include RTP \cite{rfc:3550}, one
 of the most widely deployed real-time transport protocols; the WebRTC Data
-Channel \cite{}, which tunnels peer-to-peer SCTP associations over a DTLS
-association over UDP; and QUIC \cite{}, which provides a modern alternative 
+Channel \cite{draft-ietf-rtcweb-data-channel-13}, which tunnels
+peer-to-peer SCTP associations over a DTLS association over UDP; and QUIC
+\cite{draft-tsvwg-quic-protocol-02}, which provides a modern alternative 
 to TCP, implemented over UDP.
@@
 Evidence that these services can be deployed above UDP exists in the form
-of the WebRTC data channel \cite{}.  
+of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13}.  
 \todo{expand}
@@
 Hollywood protocol, and earlier measurements by Honda \emph{et al}.\
-\cite{} also suggest that deployment over TCP is possible.
+\cite{honda:2011:extend-tcp} also suggest that deployment over TCP is
+possible.
 

### 2016-05-23 paper.tex
Completes Partial Reliability to-dos

@@
 is injecting data into an existing TCP flow. As a result, our connection
-may be disrupted. Honda et al. \cite{honda:2011:extend-tcp} conducted
+may be disrupted. Honda \emph{et al}.\ \cite{honda:2011:extend-tcp} conducted
 experiments across 135 paths on the Internet, to determine support for
-inconsistent retransmissions. The observed that the majority of paths
+inconsistent retransmissions. They observed that the majority of paths
 delivered inconsistent retransmissions successfully. On Port 80 (HTTP), the
@@
 We conducted further deployment experiments using
-inconsistent retransmissions \cite{mcquistin2016hollywood}, and found that
-fixed-line networks delivered inconsistent retransmissions successfully,
-but that delivery of the original segment was common on cellular networks.
-\todo{expand this}
+inconsistent retransmissions \cite{mcquistin2016hollywood}, testing all 
+major UK providers. We found that 100\% of tested fixed-line networks
+delivered inconsistent retransmissions successfully. However, the delivery
+of the original segment is common on cellular networks, with only 25\% of
+tested networks delivering inconsistent retransmissions successfully and
+reliably. The behaviour observed when evaluating cellular networks was
+consistent with that of a transparent, split-connection TCP cache. Segments
+were lost, but were retransmitted (with the IP address of the sender) by a
+middlebox in the network. It is likely that these caches are deployed close
+to the wireless link, given its relatively high rate of non-congestive
+loss.
 
@@
 
-\todo{Add discussion on checksumming to maintain message integrity}
+The use of inconsistent retransmissions may interact negatively with
+caching and resegmenting middleboxes, resulting in the corruption of
+messages between sender and receiver. A message may be formed from the
+original message, and an inconsistent retransmission, given how the
+mechanism uses the TCP sequence space. To protect against this, a checksum
+must be attached to each message, to allow the receiver to verify its
+integrity. The role of a checksum may also be fulfilled by using a secure
+transport protocol, such as DTLS \cite{rfc:6347}.
 

### 2016-05-23 paper.tex
Adds Transport Service and Return Value to API table

@@
 \usepackage{balance}
+\usepackage{tabularx}
 
@@
 
-\begin{table}
+\begin{table*}
   \centering
-  \begin{tabular}{ll}
+  \begin{tabularx}{\textwidth}{llll}
     \toprule
-      Function & Parameters \\
+      Transport Service            & Function & Parameters                                     & Return Value \\
     \midrule
-      \texttt{socket} & \texttt{af} -- Address family \\
-                      & \texttt{st} -- Socket type \\
-      \texttt{close} & \texttt{sd} -- Socket descriptor \\
+                                   & \texttt{socket}         
+                                              & \texttt{af} -- Address family                  & Socket descriptor \\
+                                   &          & \texttt{st} -- Socket type & \\
+                                   & \texttt{close}          
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
     \midrule
-      \texttt{bind} & \texttt{sd} -- Socket descriptor \\
-                    & \texttt{addr} -- Address to bind to \\
-                    & \texttt{addrlen} -- Length of \texttt{addr} \\
-      \texttt{listen} & \texttt{sd} -- Socket descriptor \\
-            \texttt{accept} & \texttt{sd} -- Socket descriptor \\
-                    & \texttt{addr} -- Address of peer \\
-                    & \texttt{addrlen} -- Length of \texttt{addr} \\
-      \texttt{connect} & \texttt{addr} -- Address to connect to \\
-                    & \texttt{addrlen} -- Length of \texttt{addr} \\
+      \textbf{Connection oriented} & \texttt{bind}           
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   &          & \texttt{addr} -- Address to bind to & \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
+                                   & \texttt{listen}         
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   & \texttt{accept}        
+                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
+                                   &          & \texttt{addr} -- Address of peer & \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
+                                   & \texttt{connect}        
+                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
     \midrule
-      \texttt{set\_po\_delay} & \texttt{delay} -- Playout delay (in ms) \\
+      \textbf{Deadlines}           & \texttt{set\_po\_delay} 
+                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
     \midrule
-      \texttt{send\_message} & \texttt{sd} -- Socket descriptor \\
-      & \texttt{buf} -- Message data \\
-      & \texttt{len} -- Length of message data \\
-      & \texttt{seq\_num} -- Sequence number \\
-      & \texttt{deadline} -- Relative deadline of message (in ms) \\ 
-      & \texttt{depends\_on} -- \texttt{seq\_num} of dependency \\
-      & \texttt{substream} -- Substream identifier \\
-      \texttt{recv\_message} & \texttt{sd} -- Socket descriptor \\
-      & \texttt{buf} -- Buffer for message data \\
-      & \texttt{len} -- Size of \texttt{buf} \\
-      & \texttt{substream} -- Substream identifier \\
-      & \todo{return value, rather than param?}\\
+      \textbf{Message oriented}    & \texttt{send\_message}  
+                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
+                                   &          & \texttt{buf} -- Message data & \\
+                                   &          & \texttt{len} -- Length of message data & \\
+                                   &          & \texttt{seq\_num} -- Sequence number & \\
+      \emph{Deadlines}             &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\ 
+      \emph{Dependencies}          &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
+      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
+                                   & \texttt{recv\_message}  
+                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
+                                   &          & \texttt{buf} -- Buffer for message data & \\
+                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
+      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
+                                   &          & \todo{return value, rather than param?} & \\
     \bottomrule
-  \end{tabular}
-  \caption{Outline transport API for real-time applications 
-          \todo{show return values of these calls}}
+  \end{tabularx}
+  \caption{Outline transport API for real-time applications. Return values
+  shown are for successful calls; in all cases, -1 is returned in the event
+  of an error}
   \label{tab:api}
-\end{table}
+\end{table*}
 

### 2016-05-23 paper.tex
Adds TCP Hollywood architecture diagram, and accompanying text

@@
 
+The architecture of TCP Hollywood \cite{mcquistin2016hollywood} is shown in
+Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
+services described in Section \ref{sec:services}, with functionality split
+across an intermediary layer (in userspace), and a set of modifications to
+the kernel. The intermediary layer is responsible for providing the
+message-oriented abstraction, and exposing the API for partial reliability.
+The kernel modifications implement the deadline and dependency services,
+as well as enabling out-of-order delivery of incoming segments. This split
+allows applications to use a single API, whether or not the kernel
+modifications are available. TCP Hollywood inherits standard TCP's
+congestion control mechanism. Clearly, this loss-based mechanism is not
+suitable for real-time applications, and a delay-based algorithm should
+be adopted.
+
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
+ \caption{TCP Hollywood architecture}
+\label{diagram:tcp-hollywood}
+\end{figure}
+
 Taken together, the wide experiences with the WebRTC Data Channel and QUIC

### 2016-05-23 paper.tex
Adds draft abstract

@@
 \begin{document}
-\title{Implementing Real-time Transport Services over an Ossified Network}
+\title{Implementing Real-Time Transport Services over an Ossified Network}
 \numberofauthors{3}
@@
 %  - Say what follows from your solution
+
+Real-time applications require a set of transport services not currently
+provided by widely-deployed transport protocols. Ossification prevents the
+deployment of a novel protocols, restricting solutions to protocols using
+either TCP or UDP as a substrate. We describe the transport services
+required by real-time applications. We show that, in the short term (i.e.,
+while UDP is blocked at current levels), TCP offers a feasible substrate
+for providing these services. Over the longer term, protocols using UDP
+may reduce the number of networks blocking UDP, enabling a shift towards
+its use as a demultiplexing layer for novel transport protocols.
 

### 2016-05-23 paper.tex
switch to 10pt + sig-alternet edit for compability with microtype

@@
-\documentclass{sig-alternate-05-2015}
+\documentclass[10pt]{sig-alternate-05-2015}
 \usepackage[l2tabu,orthodox]{nag}
@@
 (1) develop a taxonomy of \emph{transport services}, that is, to identify the
-features that comprise, and can be combined, to form complete transport
+features that comprise, and can be combined to form, complete transport
 protocols; and (2) to develop an abstract API for applications to request
@@
 The presence of deadlines, in a network that offers only a best-effort
-packet deliver service, implies that the transport must provide
+packet delivery service, implies that the transport must provide
 \emph{partial reliability}.
@@
 Given deadlines and partial reliability, the next important transport
-service is \emph{dependency management}. There is no point sending data
-that cannot be used, because it depends on previous data that was lost. 
-\todo{...complex trade-off, since you might want to send data that misses its 
-   deadline, if it allows later data to be used.}
+service is \emph{dependency management}. There is no point in sending data
+that cannot be used because it depends on previous data that was not
+received. Providing this service is complicated by the two ways in which
+data can be \emph{useful} to applications: it may itself by played out, or
+it may be needed as part of the application's decoding chain. For example,
+an I frame provides utility both in being played out, and in allowing for
+subsequent P and B frames. As a result, dependency information may result
+in deadline information being overruled: data may be sent even if it cannot
+be played out, if it is necessary for the decoding of other data.
 
-\todo{...following from this, we have the concept of application-level framing
-   to make the best use of data, irrespective of losses. This implies that
-   the transport provides a \emph{message oriented} service that respects
-   ADU boundaries.}
+Providing partial reliability, through deadlines and dependencies, requires
+the use of application-level framing \cite{clark:1990:architecture} to
+make the best use of payload data, irrespective of packet loss. At the
+transport layer, this implies that we provide a \emph{message oriented}
+service, that maintains ADU boundaries.
 
-\todo{...we note that many multimedia applications make use of multiple streams. 
-   For example, audio and video are logically separate, etc.
-   can be supported by opening multiple transport connections, but
-   desirable to support a \emph{sub-stream} service to reduce overheads.}
+Many multimedia applications make use of multiple data flows. For example,
+a simple IPTV application will maintain separate audio and video flows.
+These could be sent across multiple transport-layer connections, but
+overheads can be reduced by multiplexing these flows on a single
+connection. A \emph{sub-stream} service allows this.
 
-\todo{...finally, we note that \emph{congestion control} is important. 
-   motivate: historically there has been a belief that real-time doesn't 
-   do congestion control, needs an isochronous channel. 
-   not practical for the Internet. 
-   Also, experience has shown not necessary for many real-time applications.
-   Yes, there are some that need constant bandwidth and cannot adapt, but
-   many are adaptive. 
-   Better user experience to adapt to fit the channel, rather than recover
-   from loss after the channel discards packets that don't fit. }
+We note that \emph{congestion control} is important. Clearly, the historic
+belief that real-time applications require an isochronous channel, and
+cannot be subjected to congestion control, is not practical on the
+Internet. Further, while there is a number of applications that are not
+adaptive (and require constant bandwidth), an increasing number are
+adaptive, and would better serve users by adapting to the available
+bandwidth.
 
@@
     \midrule
-      \textbf{Connection oriented} & \texttt{bind}           
+      Connection oriented          & \texttt{bind}
                                               & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
@@
     \midrule
-      \textbf{Deadlines}           & \texttt{set\_po\_delay} 
+      Deadlines                    & \texttt{set\_po\_delay}
                                               & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
     \midrule
-      \textbf{Message oriented}    & \texttt{send\_message}  
+      Message oriented             & \texttt{send\_message}
                                               & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
@@
                                    &          & \texttt{seq\_num} -- Sequence number & \\
-      \emph{Deadlines}             &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\ 
-      \emph{Dependencies}          &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
-      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
+      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
+      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
+      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                    & \texttt{recv\_message}
@@
                                    &          & \texttt{len} -- Size of \texttt{buf} & \\
-      \emph{Sub-streams}           &          & \texttt{substream} -- Substream identifier & \\
+      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
                                    &          & \todo{return value, rather than param?} & \\
@@
 Evidence that these services can be deployed above UDP exists in the form
-of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13}.  
-\todo{expand}
-
-\todo{also QUIC}
+of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13} and
+QUIC protocol \cite{draft-tsvwg-quic-protocol-02}.
+The former is a peer-to-peer protocol, comprising an SCTP association
+running over DTLS, itself running over a UDP flow negotiated via an SDP
+\cite{RFC4566} offer/answer exchange \cite{RFC3264} as part of a WebRTC
+session \cite{jennings:2013:rtcweb} (the WebRTC media channel uses RTP
+running over UDP also, further showing the utility of UDP-based data).
+This has been deployed in popular web browsers, with global deployment, and
+demonstrated to be effective.  The latter is implemented by Google in their
+Chrome browser, and used as an alternative to TCP has a significant
+fraction of web traffic downloads from their domain.
 
@@
 Section \ref{sec:ossification}, there are also reasons for providing these
-services running over TCP. 
-\todo{
-  these services are deployed over TCP now, but have an inconvenient API that
-  imposes lots of work on application developers, and with higher than desired
-  latency. 
-  TCP Hollywood offers a way to fix both issues.
-  Brief summary.
-  Obviously we need to change TCP congestion control too.
-}
+services over TCP, since there are a significant fraction of networks that
+block UDP traffic. It is clearly possible to run real-time traffic over
+TCP, as demonstrated by applications such as NetFlix or the BBC iPlayer
+that comprise the majority of Internet traffic. However, TCP has a
+inconvenient API that imposes lots of work on application developers, and
+introduces higher than desired latency. We have shown how to address these
+issues, and provide the full set of transport services we propose in
+Section \ref{sec:services} in previous work, with our TCP Hollywood
+proposal \cite{mcquistin2016hollywood}.
 
-The architecture of TCP Hollywood \cite{mcquistin2016hollywood} is shown in
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
+ \caption{TCP Hollywood architecture}
+\label{diagram:tcp-hollywood}
+\end{figure}
+
+The architecture of TCP Hollywood is shown in
 Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
 services described in Section \ref{sec:services}, with functionality split
-across an intermediary layer (in userspace), and a set of modifications to
+across an intermediary layer (in user-space), and a set of modifications to
 the kernel. The intermediary layer is responsible for providing the
@@
 
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
- \caption{TCP Hollywood architecture}
-\label{diagram:tcp-hollywood}
-\end{figure}
+\todo{Stephen: give more detail, and expand this to a whole column}
+
+\todo{discuss how we need to change TCP congestion control}
 
@@
 %==================================================================================================
-\section{Acknowledgements}
+%\section{Acknowledgements}
 

### 2016-05-23 paper.tex
minor corrections

@@
     The \texttt{send\_message()} call takes four additional parameters
-    compared to the standard Berkeley Sockets API. These are 1) a message
-    sequence number, that can be used to re-order messages and detect
-    message loss; 2) a relative deadline, which is combined with an
-    estimate of the current round-trip-time, and the time that the message
-    has spent in the sending buffer, to determine if a message will arrive
-    in time to be played-out; 3) the sequence number of any message on
-    which this depends, for example the sequence number of a video I-frame
-    on which a P-frame is predicted; and 4) a sub-stream identifier, used,
-    for example, to differentiate audio, video, sub-title, control, and
-    repair streams.
+    compared to the standard Berkeley Sockets API. These are 1) a
+    message sequence number, that can be used to re-order messages and
+    detect message loss; 2) a relative deadline, which is combined
+    with an estimate of the current round-trip-time, and the time that
+    the message has spent in the sending buffer, to determine if a
+    message will arrive in time to be played-out; 3) the message
+    sequence number of any message on which this depends, for example,
+    of a video I-frame on which a P-frame is predicted; and 4) a
+    sub-stream identifier, used, for example, to differentiate audio,
+    video, sub-title, control, and repair streams.
 
@@
 A message that won't arrive within its lifetime is considered to have
-\emph{expired}. Applications specify the application-level sequence number
-of a message they depend on (\texttt{depends\_on}). If the message depends
-on a message that has expired, then it too is considered to have expired.
-The partial reliability service follows from the deadline and dependency
-services: messages will be reliably transmitted until they expire.
+\emph{expired}. A message is also considered to have expired if its
+message sequence number dependency, \texttt{depends\_on},
+%on which a message depends, . If the message depends on a message that has expired, then it too is considered to have
+has expired. A partial reliability service follows
+from this deadline and dependency service: messages will be reliably
+transmitted until they expire.
 

### 2016-05-23 paper.tex
Tweaks API table

--- 2016-05-23 paper.tex
@@
-\documentclass[10pt]{sig-alternate-05-2015}
+\documentclass{sig-alternate-05-2015}
 \usepackage[l2tabu,orthodox]{nag}
@@
 Evidence that these services can be deployed above UDP exists in the form
-of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13} and
-QUIC protocol \cite{draft-tsvwg-quic-protocol-02}.
-The former is a peer-to-peer protocol, comprising an SCTP association
-running over DTLS, itself running over a UDP flow negotiated via an SDP
-\cite{RFC4566} offer/answer exchange \cite{RFC3264} as part of a WebRTC
-session \cite{jennings:2013:rtcweb} (the WebRTC media channel uses RTP
-running over UDP also, further showing the utility of UDP-based data).
-This has been deployed in popular web browsers, with global deployment, and
-demonstrated to be effective.  The latter is implemented by Google in their
-Chrome browser, and used as an alternative to TCP has a significant
-fraction of web traffic downloads from their domain.
+of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13}.  
+\todo{expand}
+
+\todo{also QUIC}
 
@@
 Section \ref{sec:ossification}, there are also reasons for providing these 
-services over TCP, since there are a significant fraction of networks that
-block UDP traffic. It is clearly possible to run real-time traffic over
-TCP, as demonstrated by applications such as NetFlix or the BBC iPlayer
-that comprise the majority of Internet traffic. However, TCP has a
-inconvenient API that imposes lots of work on application developers, and
-introduces higher than desired latency. We have shown how to address these
-issues, and provide the full set of transport services we propose in
-Section \ref{sec:services} in previous work, with our TCP Hollywood
-proposal \cite{mcquistin2016hollywood}.
+services running over TCP. 
+\todo{
+  these services are deployed over TCP now, but have an inconvenient API that
+  imposes lots of work on application developers, and with higher than desired
+  latency. 
+  TCP Hollywood offers a way to fix both issues.
+  Brief summary.
+  Obviously we need to change TCP congestion control too.
+}
 
-\begin{figure}[t]
- \centering
- \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
- \caption{TCP Hollywood architecture}
-\label{diagram:tcp-hollywood}
-\end{figure}
-
-The architecture of TCP Hollywood is shown in
+The architecture of TCP Hollywood \cite{mcquistin2016hollywood} is shown in
 Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
 services described in Section \ref{sec:services}, with functionality split
-across an intermediary layer (in user-space), and a set of modifications to
+across an intermediary layer (in userspace), and a set of modifications to
 the kernel. The intermediary layer is responsible for providing the
@@
 
-\todo{Stephen: give more detail, and expand this to a whole column}
-
-\todo{discuss how we need to change TCP congestion control}
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
+ \caption{TCP Hollywood architecture}
+\label{diagram:tcp-hollywood}
+\end{figure}
 
@@
 %==================================================================================================
-%\section{Acknowledgements}
+\section{Acknowledgements}
 

### 2016-05-23 paper.tex
intro edits and comment

--- 2016-05-23 paper.tex
@@
 
-Real-time applications are increasingly important in the Internet.
-We want to make it easier to write these applications, while improving
-the quality of experience for users of those applications by lowering
-latency and increasing robustness.
-In this, we are challenged by the limitations of the standard Internet
-transport protocols, and by the ossified nature of the network, which makes
+Real-time applications are increasingly
+%important
+present in the Internet.
+%We want to make it easier to write these applications
+Such applications should be easier to program, while also improving
+the quality of experience for users.
+ %of those applications
+by lowering latency and increasing robustness.
+%In this, we are
+This limitations of standard Internet transport protocols makes this a
+challenging target. Moreover, the ossified nature of the network makes
 it increasingly difficult to deploy new transports.
@@
 
-In practice, the only protocols that can be widely used in the Internet are
-TCP and UDP. Other protocols are blocked by firewalls and other middleboxes.
-TCP provides sophisticated congestion control, coupled with a reliable,
-ordered, byte stream API. These are suitable for many applications, but are
-not necessarily appropriate for real-time traffic.  UDP exposes the best-effort
-IP packet delivery service, offering flexibility to develop new protocols,
-but at the cost of requiring completely new protocol mechanisms to be
-defined. Both protocols are usable for real-time applications, but neither
-really provides the right services and API.  This forces each application
-to re-invent mechanisms that should be provided by the transport, increasing
-costs and complexity, raising barriers to innovation.
+In practice, only UDP and TCP are widely usable in the Internet,
+since remaining protocols are blocked by firewalls and other
+middleboxes. UDP exposes the best-effort IP packet delivery service,
+offering the flexibility to develop new protocols. The high costs
+associated involve defining completely new protocol mechanisms. In
+contrast TCP mechanisms are well defined, consisting of sophisticated
+congestion control coupled with a reliable, ordered, byte stream API.
+These are proven suitable for many applications, but are inappropriate
+for real-time traffic. While both protocols may be used for real-time
+applications, neither really provides the right services and API. This
+forces each application to re-invent or re-interpret mechanisms that
+should be provided by the transport. The increased costs and
+complexity of doing so raise barriers to innovation.
 
@@
 
-In this paper we discuss the appropriate set of transport services and APIs
-for real-time applications.
-We demonstrate that it is possible to realise these services and APIs
-in the context of both TCP and UDP, despite the limitations imposed by the
-legacy protocols, middleboxes, and ossification of the network.
-We present initial results to show that, despite the observed ossification,
-the network has the flexibility to deploy new transport protocols, if care
-is taken to reinterpret layer boundaries and deploy those new transports in
-the context of TCP and UDP.
+In this paper we identify and present the appropriate set of transport
+services and APIs for real-time applications, and demonstrate their
+merit by implementing a proof-of-concept. We show that it is possible
+to realise real-time services and APIs in the context of both TCP and
+UDP, despite the limitations imposed by their legacies, and by
+middleboxes, and ossification of the network.  Initial experiments
+with our implementation suggest that
+%to show that, despite network ossification,
+the network has the flexibility to deploy new transport protocols,
+provided care is taken to reinterpret layer boundaries that respect
+conventional UDP and TCP.\todo{I can infer the meaning of previous sentence having worked on the problem; otherwise may be very ambiguous to the reader.}
+%and deploy those new transports in the context of TCP and UDP.
 
@@
 
-Our contributions are to identify the needs of real-time applications, and
-the appropriate transport services and APIs to support those needs;
-to illustrate how those transport services can be realised on the current
-Internet, in the context of UDP and TCP deployments; and
-to present some initial measurement results to show that the mechanisms we
-propose ought to be usable in the public Internet.
+% Our contributions
+In doing so we make three main contributions. First, we make explicit
+the needs of real-time applications, as well as the appropriate
+transport services and APIs to support those needs. Second, we
+illustrate an example realisation of those transport services on the
+current Internet, in the context of UDP and TCP deployments. Finally,
+we present initial measurement results that suggest the proposed
+mechanisms ought to be usable in the public Internet.
 
@@
 
-We begin, in Section \ref{sec:services} by discussing transport services
-for real-time applications, and outlining the common conceptual API that
-those applications use.
-This is followed, in Section \ref{sec:ossification}, by a review of the
-deployment considerations for new protocols, caused by ossification of
-the network, with Section \ref{sec:partial} considering, in particular, how
-TCP reliability semantics can evolve within the constraints of the deployed
-infrastructure.
-Section \ref{sec:realising} outlines how the transport services we have
-identified can be realised in practical networks.
+We begin in Section \ref{sec:services} by discussing transport
+services for real-time applications, and outlining the common
+conceptual API that those applications use. This is followed in
+Section \ref{sec:ossification} by a review of deployment
+considerations for new protocols, caused by ossification of the
+network. Section \ref{sec:partial} considers, in particular, how TCP
+reliability semantics can evolve within the constraints of the
+existing infrastructure. The semantics are realised and put into
+practice in Section \ref{sec:realising}.
+% outlines how the transport services we have identified can be
+% realised in practical networks.
 Finally, Section \ref{sec:related} discusses related work, and Section

### 2016-05-23 paper.tex
Work on Section 5

--- 2016-05-23 paper.tex
@@
 Evidence that these services can be deployed above UDP exists in the form
-of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13}.  
-\todo{expand}
-
-\todo{also QUIC}
+of the WebRTC data channel \cite{draft-ietf-rtcweb-data-channel-13} and
+QUIC protocol \cite{draft-tsvwg-quic-protocol-02}.  
+The former is a peer-to-peer protocol, comprising an SCTP association
+running over DTLS, itself running over a UDP flow negotiated via an SDP
+\cite{RFC4566} offer/answer exchange \cite{RFC3264} as part of a WebRTC
+session \cite{jennings:2013:rtcweb} (the WebRTC media channel uses RTP
+running over UDP also, further showing the utility of UDP-based data). 
+This has been deployed in popular web browsers, with global deployment, and
+demonstrated to be effective.  The latter is implemented by Google in their
+Chrome browser, and used as an alternative to TCP has a significant
+fraction of web traffic downloads from their domain.
 
@@
 Section \ref{sec:ossification}, there are also reasons for providing these 
-services running over TCP. 
-\todo{
-  these services are deployed over TCP now, but have an inconvenient API that
-  imposes lots of work on application developers, and with higher than desired
-  latency. 
-  TCP Hollywood offers a way to fix both issues.
-  Brief summary.
-  Obviously we need to change TCP congestion control too.
-}
+services over TCP, since there are a significant fraction of networks that
+block UDP traffic. It is clearly possible to run real-time traffic over
+TCP, as demonstrated by applications such as NetFlix or the BBC iPlayer
+that comprise the majority of Internet traffic. However, TCP has a
+inconvenient API that imposes lots of work on application developers, and
+introduces higher than desired latency. We have shown how to address these
+issues, and provide the full set of transport services we propose in
+Section \ref{sec:services} in previous work, with our TCP Hollywood
+proposal \cite{mcquistin2016hollywood}.
 
-The architecture of TCP Hollywood \cite{mcquistin2016hollywood} is shown in
+\begin{figure}[t]
+ \centering
+ \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
+ \caption{TCP Hollywood architecture}
+\label{diagram:tcp-hollywood}
+\end{figure}
+
+The architecture of TCP Hollywood is shown in
 Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
 services described in Section \ref{sec:services}, with functionality split
-across an intermediary layer (in userspace), and a set of modifications to
+across an intermediary layer (in user-space), and a set of modifications to
 the kernel. The intermediary layer is responsible for providing the

### 2016-05-23 paper.tex
Adds paragraph on HAS protocols to Section 3

--- 2016-05-23 paper.tex
@@
 \usepackage{tabularx}
+
+\newcommand*\rot{\rotatebox{90}}
 
@@
                                               & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
-                                   &          & \texttt{buf} -- Buffer for message data & \\
+                                   &          & \texttt{buf} -- Buffer for message data        & Substream identifier \\
                                    &          & \texttt{len} -- Size of \texttt{buf} & \\
-      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
-                                   &          & \todo{return value, rather than param?} & \\
     \bottomrule
@@
 
-\todo{outline why this is not true in practice}
+There are performance and security benefits that can be attained by
+adding transport-layer functionality \emph{within} the network. For
+example, a firewall can better protect the network if it can detect payload
+anomalies.
 
@@
 
-\todo{Other motivation for using TCP: HTTP-based streaming protocols}
+Beyond the availability of UDP, it is often necessary to use TCP because
+HTTP is being used at the application-layer. For real-time systems, this
+is likely to be an HTTP adaptive streaming (HAS) protocol, such as
+MPEG-DASH or Apple's HLS. Using TCP as a substrate enables the use of these
+protocols, allowing applications to benefit from the existing
+infrastructure that supports them.
 
@@
 supports such an abstraction for datagram protocols.
-
-\todo{discuss how the API can be modified to accept/delivery a sequence
-      of messages, rather than a byte stream, and how with consistent
-      segmentation at the sender and COBS framing, this can be efficient}
 

### 2016-05-23 paper.tex

@@
 
-Real-time applications are increasingly important in the Internet. 
-We want to make it easier to write these applications, while improving
-the quality of experience for users of those applications by lowering
-latency and increasing robustness. 
-In this, we are challenged by the limitations of the standard Internet 
-transport protocols, and by the ossified nature of the network, which makes
+Real-time applications are increasingly
+%important
+present in the Internet.
+%We want to make it easier to write these applications
+Such applications should be easier to program, while also improving
+the quality of experience for users.
+ %of those applications
+by lowering latency and increasing robustness.
+%In this, we are
+This limitations of standard Internet transport protocols makes this a
+challenging target. Moreover, the ossified nature of the network makes
 it increasingly difficult to deploy new transports.
@@
 
-In practice, the only protocols that can be widely used in the Internet are
-TCP and UDP. Other protocols are blocked by firewalls and other middleboxes.  
-TCP provides sophisticated congestion control, coupled with a reliable,
-ordered, byte stream API. These are suitable for many applications, but are
-not necessarily appropriate for real-time traffic.  UDP exposes the best-effort 
-IP packet delivery service, offering flexibility to develop new protocols,
-but at the cost of requiring completely new protocol mechanisms to be
-defined. Both protocols are usable for real-time applications, but neither
-really provides the right services and API.  This forces each application
-to re-invent mechanisms that should be provided by the transport, increasing 
-costs and complexity, raising barriers to innovation.
+In practice, only UDP and TCP are widely usable in the Internet,
+since remaining protocols are blocked by firewalls and other
+middleboxes. UDP exposes the best-effort IP packet delivery service,
+offering the flexibility to develop new protocols. The high costs
+associated involve defining completely new protocol mechanisms. In
+contrast TCP mechanisms are well defined, consisting of sophisticated
+congestion control coupled with a reliable, ordered, byte stream API.
+These are proven suitable for many applications, but are inappropriate
+for real-time traffic. While both protocols may be used for real-time
+applications, neither really provides the right services and API. This
+forces each application to re-invent or re-interpret mechanisms that
+should be provided by the transport. The increased costs and
+complexity of doing so raise barriers to innovation.
 
@@
 
-In this paper we discuss the appropriate set of transport services and APIs
-for real-time applications.
-We demonstrate that it is possible to realise these services and APIs
-in the context of both TCP and UDP, despite the limitations imposed by the
-legacy protocols, middleboxes, and ossification of the network. 
-We present initial results to show that, despite the observed ossification,
-the network has the flexibility to deploy new transport protocols, if care
-is taken to reinterpret layer boundaries and deploy those new transports in
-the context of TCP and UDP.
+In this paper we identify and present the appropriate set of transport
+services and APIs for real-time applications, and demonstrate their
+merit by implementing a proof-of-concept. We show that it is possible
+to realise real-time services and APIs in the context of both TCP and
+UDP, despite the limitations imposed by their legacies, and by
+middleboxes, and ossification of the network.  Initial experiments
+with our implementation suggest that
+%to show that, despite network ossification,
+the network has the flexibility to deploy new transport protocols,
+provided care is taken to reinterpret layer boundaries that respect
+conventional UDP and TCP.\todo{I can infer the meaning of previous sentence having worked on the problem; otherwise may be very ambiguous to the reader.}
+%and deploy those new transports in the context of TCP and UDP.
 
@@
 
-Our contributions are to identify the needs of real-time applications, and
-the appropriate transport services and APIs to support those needs; 
-to illustrate how those transport services can be realised on the current
-Internet, in the context of UDP and TCP deployments; and 
-to present some initial measurement results to show that the mechanisms we
-propose ought to be usable in the public Internet.
+% Our contributions
+In doing so we make three main contributions. First, we make explicit
+the needs of real-time applications, as well as the appropriate
+transport services and APIs to support those needs. Second, we
+illustrate an example realisation of those transport services on the
+current Internet, in the context of UDP and TCP deployments. Finally,
+we present initial measurement results that suggest the proposed
+mechanisms ought to be usable in the public Internet.
 
@@
 
-We begin, in Section \ref{sec:services} by discussing transport services 
-for real-time applications, and outlining the common conceptual API that
-those applications use. 
-This is followed, in Section \ref{sec:ossification}, by a review of the
-deployment considerations for new protocols, caused by ossification of
-the network, with Section \ref{sec:partial} considering, in particular, how
-TCP reliability semantics can evolve within the constraints of the deployed
-infrastructure.
-Section \ref{sec:realising} outlines how the transport services we have
-identified can be realised in practical networks.
+We begin in Section \ref{sec:services} by discussing transport
+services for real-time applications, and outlining the common
+conceptual API that those applications use. This is followed in
+Section \ref{sec:ossification} by a review of deployment
+considerations for new protocols, caused by ossification of the
+network. Section \ref{sec:partial} considers, in particular, how TCP
+reliability semantics can evolve within the constraints of the
+existing infrastructure. The semantics are realised and put into
+practice in Section \ref{sec:realising}.
+% outlines how the transport services we have identified can be
+% realised in practical networks.
 Finally, Section \ref{sec:related} discusses related work, and Section
@@
     The \texttt{send\_message()} call takes four additional parameters
-    compared to the standard Berkeley Sockets API. These are 1) a message
-    sequence number, that can be used to re-order messages and detect 
-    message loss; 2) a relative deadline, which is combined with an
-    estimate of the current round-trip-time, and the time that the message
-    has spent in the sending buffer, to determine if a message will arrive
-    in time to be played-out; 3) the sequence number of any message on
-    which this depends, for example the sequence number of a video I-frame
-    on which a P-frame is predicted; and 4) a sub-stream identifier, used,
-    for example, to differentiate audio, video, sub-title, control, and
-    repair streams.
+    compared to the standard Berkeley Sockets API. These are 1) a
+    message sequence number, that can be used to re-order messages and
+    detect message loss; 2) a relative deadline, which is combined
+    with an estimate of the current round-trip-time, and the time that
+    the message has spent in the sending buffer, to determine if a
+    message will arrive in time to be played-out; 3) the message
+    sequence number of any message on which this depends, for example,
+    of a video I-frame on which a P-frame is predicted; and 4) a
+    sub-stream identifier, used, for example, to differentiate audio,
+    video, sub-title, control, and repair streams.
 

### 2016-05-23 paper.tex
Adds updated TCP Hollywood architecture diagram
Updates architecture diagram
2.1 edits ongoing; pushing before conflicts appear.

@@
 \usepackage{tabularx}
+\usepackage{xspace}
 
+\newcommand{\ie}{{i.e.}\xspace}
 \newcommand*\rot{\rotatebox{90}}
@@
 
-The work in TAPS is valuable, because it gives us a vocabulary for
-discussing the components of transport protocols.  We can then use that
-vocabulary when discussing the needs of real-time applications, and the
-protocols needed to support those needs.
+The work in TAPS provides a vocabulary for discussing the components of
+transport protocols.  The vocabulary is useful when discussing the needs of
+real-time applications, and the protocols to support them.
 
-The essential feature of real-time applications is clearly timing. That is,
-since these applications convey real-time data, they all have some concept
-of a \emph{deadline} by which that data must be presented, after which it
-becomes useless.
-The deadline might be short or long. Interactive applications, such as
-telephony, video conferencing, or telepresence, require low end-to-end
-latency, with the deadline for presenting the media (playing the audio,
-displaying the video frame) being tens, or perhaps low hundreds, of
-milliseconds after it was captured. Non-interactive applications, such as
-broadcast TV, may have deadlines in the order of seconds, or even tens of
-seconds for on-demand programming.
-The deadlines in many networked multimedia applications are unusual when
-compared to other real-time systems. They are simultaneously flexible and
-strict: flexible in that the exact value of the deadline is typically not
-important, provided it is of the right order-of-magnitude for the
-application, but strict in that any particular deadline provides a hard
-cut-off, after which the data is useless.
+Timing is the most salient feature of real-time applications. Since their data
+must be conveyed with real-time demands, they all have some concept of a
+\emph{deadline}. Data that fails to present within the deadline is otherwise
+useless.
+% Deadlines may be short or long.
+The `slack' in a deadline depends on the application. Interactive applications,
+such as telephony, video conferencing, or telepresence, require low end-to-end
+latency. Their deadlines for presenting the media, \ie playing the audio and
+displaying the video frame, range from tens to a few hundred milliseconds.
+Non-interactive application deadlines associated with broadcast and on-demand
+programming are on the order of seconds.
+% , such as broadcast TV, may have deadlines on the order of seconds,
+% or even tens of seconds for on-demand programming.
+
+Networked multimedia deadlines are unusual when compared to other real-time
+systems. They are simultaneously flexible and strict: flexible in that the exact
+value of the deadline is typically not important, provided it is of the right
+order-of-magnitude for the application, but strict in that any particular
+deadline provides a hard cut-off, after which the data is useless. \todo{this
+para is largely unconvincing -- flexible is fine; strict is less believable, \ie
+how are RT systems any different when the deadline expires? (Many RTSs will
+discard, too.)}
 
@@
  \centering
- \includegraphics[scale=0.965]{figures/tcp-hollywood.pdf}
- \caption{TCP Hollywood architecture}
+ \includegraphics[scale=0.45]{figures/tcp-hollywood.pdf}
+ \caption{TCP Hollywood architecture \todo{too similar to NOSSDAV fig 1?}}
 \label{diagram:tcp-hollywood}
@@
 
-The architecture of TCP Hollywood is shown in
-Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
-services described in Section \ref{sec:services}, with functionality split
-across an intermediary layer (in user-space), and a set of modifications to
-the kernel. The intermediary layer is responsible for providing the
-message-oriented abstraction, and exposing the API for partial reliability.
-The kernel modifications implement the deadline and dependency services,
-as well as enabling out-of-order delivery of incoming segments. This split
-allows applications to use a single API, whether or not the kernel
-modifications are available. TCP Hollywood inherits standard TCP's
-congestion control mechanism. Clearly, this loss-based mechanism is not
-suitable for real-time applications, and a delay-based algorithm should
-be adopted.
+The architecture of TCP Hollywood is shown in Figure
+\ref{diagram:tcp-hollywood}. TCP Hollywood implements all of the services
+described in Section \ref{sec:services}, splitting functionality across
+an intermediary layer in user-space, and a set of modifications to the
+kernel. This split allows applications to program against one API, whether
+or not the kernel modifications are available: the intermediary layer
+functions in both cases.
 
-\todo{Stephen: give more detail, and expand this to a whole column}
+At the sender, applications pass messages (using an API similar to that
+given in Table \ref{tab:api}) to the intermediary layer, with their
+metadata, including deadline and dependency information. At the 
+intermediary layer, COBS encoding \cite{CB97COBS} is used to escape all
+zero bytes in the message data, allowing them to be used as framing
+markers. The message's metadata is then attached to the encoded and framed
+message, before being passed to the kernel using the standard Berkeley
+sockets API.
 
+At the kernel, the message data is queued in TCP's sending buffer, while
+the metadata is held in a separate structure. Nagle's algorithm, designed
+to coalesce smaller writes into larger segments, is disabled to minimise
+latency. As segments are (re-)transmitted, their deadlines and dependencies
+are checked to ensure that the message will be useful on arrival. In the
+current version of TCP Hollywood, the dependency check does not overrule
+the deadline check: only data that can be played out will be sent. If the
+message does not pass the liveness check, the next message in the queue
+that is live will be sent instead. If this is a retransmission, then
+inconsistent retransmissions will be used: the replacement message will be
+sent with the same TCP sequence number as the original.
+
+At the receiver, segments are passed to the kernel, where they are
+initially processed as under standard TCP: duplicate acknowledgements are
+generated for out-of-order segments, for example. After this, a metadata
+entry is created, and placed in FIFO queue. When the intermediary layer
+reads from the socket, it receives the segment associated with the
+metadata entry at the head of the queue, with its TCP sequence number
+attached.
+
+At the intermediary layer on the receiver, incoming segments are scanned
+for complete messages (i.e., data between two zero bytes), which are
+decoded and passed to the application. Any message fragments are buffered,
+alongside their TCP sequence number, awaiting the arrival of the remainder
+of the message. Once the message has been reassembled, it is decoded, and
+delivered to the application.
+
+TCP Hollywood inherits standard TCP's congestion control mechanism.
+Clearly, being a loss-based algorithm, this is not appropriate for
+real-time applications.
 \todo{discuss how we need to change TCP congestion control}
@@
 message-oriented service, but allows applications to attach a time-line
-to messages. Messages are (re-)transmitted asunder standard TCP within
+to messages. Messages are (re-)transmitted as under standard TCP within
 their time-line, after which they are discarded. The mechanism by which

### 2016-05-23 paper.tex
Expands discussion of TCP Hollywood

@@
 \usepackage{tabularx}
-\usepackage{xspace}
 
-\newcommand{\ie}{{i.e.}\xspace}
 \newcommand*\rot{\rotatebox{90}}
@@
 
-The work in TAPS provides a vocabulary for discussing the components of
-transport protocols.  The vocabulary is useful when discussing the needs of
-real-time applications, and the protocols to support them.
+The work in TAPS is valuable, because it gives us a vocabulary for
+discussing the components of transport protocols.  We can then use that
+vocabulary when discussing the needs of real-time applications, and the
+protocols needed to support those needs.
 
-Timing is the most salient feature of real-time applications. Since their data
-must be conveyed with real-time demands, they all have some concept of a
-\emph{deadline}. Data that fails to present within the deadline is otherwise
-useless.
-% Deadlines may be short or long.
-The `slack' in a deadline depends on the application. Interactive applications,
-such as telephony, video conferencing, or telepresence, require low end-to-end
-latency. Their deadlines for presenting the media, \ie playing the audio and
-displaying the video frame, range from tens to a few hundred milliseconds.
-Non-interactive application deadlines associated with broadcast and on-demand
-programming are on the order of seconds.
-% , such as broadcast TV, may have deadlines on the order of seconds,
-% or even tens of seconds for on-demand programming.
-
-Networked multimedia deadlines are unusual when compared to other real-time
-systems. They are simultaneously flexible and strict: flexible in that the exact
-value of the deadline is typically not important, provided it is of the right
-order-of-magnitude for the application, but strict in that any particular
-deadline provides a hard cut-off, after which the data is useless. \todo{this
-para is largely unconvincing -- flexible is fine; strict is less believable, \ie
-how are RT systems any different when the deadline expires? (Many RTSs will
-discard, too.)}
+The essential feature of real-time applications is clearly timing. That is,
+since these applications convey real-time data, they all have some concept
+of a \emph{deadline} by which that data must be presented, after which it
+becomes useless.
+The deadline might be short or long. Interactive applications, such as
+telephony, video conferencing, or telepresence, require low end-to-end
+latency, with the deadline for presenting the media (playing the audio,
+displaying the video frame) being tens, or perhaps low hundreds, of
+milliseconds after it was captured. Non-interactive applications, such as
+broadcast TV, may have deadlines in the order of seconds, or even tens of
+seconds for on-demand programming.
+The deadlines in many networked multimedia applications are unusual when
+compared to other real-time systems. They are simultaneously flexible and
+strict: flexible in that the exact value of the deadline is typically not
+important, provided it is of the right order-of-magnitude for the
+application, but strict in that any particular deadline provides a hard
+cut-off, after which the data is useless.
 

### 2016-05-23 paper.tex
subtitles to 2.; a few substantial edits; still working my way through it

--- 2016-05-23 paper.tex
@@
 %--------------------------------------------------------------------------------------------------
-\subsection{Transport Services}
+\subsection{Timing and Transport Services}
 
@@
 order-of-magnitude for the application, but strict in that any particular
-deadline provides a hard cut-off, after which the data is useless. \todo{this
-para is largely unconvincing -- flexible is fine; strict is less believable, \ie
+deadline provides a hard cut-off, after which the data is useless.
+\todo{Is this para necessary at all?
+Its contribution is unclear and/or unconvincing -- flexible is fine; strict is less believable, \ie
 how are RT systems any different when the deadline expires? (Many RTSs will
-discard, too.)}
+discard, too. Ultimately, I get what it is trying to say; I'm less sure it needs saying) Maybe the following(?):}
+% When compared to conventional real-time systems, networked multimedia
+% deadlines are neither soft nor hard.
+It is useful to draw a distinction between in networked multimedia deadlines and
+deadlines in conventional real-time systems.
+A multimedia deadline is neither soft, since arrivals played after a deadline contribute to poor experience; nor are deadlines hard, since occasional missed messages can be tolerated.
 
-The presence of deadlines, in a network that offers only a best-effort
-packet delivery service, implies that the transport must provide
-\emph{partial reliability}.
-The network has non-zero probability of losing any particular packet.
-If forward error correction is used to repair loss, this implies there is
-always some probability that a particular packet will be non-recoverable.
-If retransmission is used to recover from losses, there is the potential
-for unbounded delay, since any retransmission may itself be lost.
-These probabilities can be estimated, to give probabilistic constraints
-on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
-the transport must support partial reliability service, accepting that
-it may not be able to deliver all data by its deadline.
-Many real-time applications run over TCP today, and TCP does not offer a
-partial reliability service. This results in occasional play-out stalls,
-when the transport blocks the application because a retransmission takes
-too long. These stalls are one of the primary causes of poor user
-experience in streaming applications, and are directly caused by the lack
-of partial reliability in TCP. Missing one frame that has not been
-delivered by its deadline is much less disruptive than a stall in play-out.
+\subsection{Partial Reliability}
+In a network that offers only a best-effort packet delivery service, deadlines
+constrain delivery to \emph{partial reliability}.
+%The network has non-zero probability of losing any particular packet.
+For example, when used to repair loss, the limits of forward error correction
+imply some probability that packet will be non-recoverable. By contrast,
+retransmissions used to recover from loss have potentially unbounded delay
+(since any retransmission may itself be lost).
+% These probabilities can be estimated, to give probabilistic constraints on
+% loss and delay, but will be non-zero. Accordingly, to meet deadlines the
+Accordingly, a transport protocol that meets deadlines should provide partial
+reliability, acknowledging that it may be unable to deliver all data by its
+deadline.
 
+Many real-time applications run over TCP today, though TCP offers no partial
+reliability service. TCP's full reliability can lead to play-out stalls
+when the application is blocked by retransmissions that take too
+long. These stalls are one of the primary causes of poor user experience in
+streaming applications.
+% , and are directly caused by the lack of partial reliability in TCP.
+For the applications under scrutiny, a missed frame that fails to deliver by its
+deadline is much less disruptive than a stall in play-out.
+
+\subsection{Message-oriented Dependencies}
 Given deadlines and partial reliability, the next important transport
@@
 
+\subsection{Connection and Congestion Control}
 We note that \emph{congestion control} is important. Clearly, the historic

### 2016-05-23 paper.tex
Adds deployability results table

--- 2016-05-23 paper.tex
@@
 \usepackage{tabularx}
+\usepackage{rotating}
+\usepackage{tikz}
 
-\newcommand*\rot{\rotatebox{90}}
+\makeatletter
+\global\let\tikz@ensure@dollar@catcode=\relax
+\makeatother
+
+\newcommand{\fail}{\tikz\draw[red,fill=red] (0,0) circle (.5ex);}
+\newcommand{\pass}{\tikz\draw[green,fill=green] (0,0) circle (.5ex);}
+\newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) circle (.5ex);}
 
@@
 inconsistent retransmissions \cite{mcquistin2016hollywood}, testing all
-major UK providers. We found that 100\% of tested fixed-line networks
+major UK providers. The results are shown in Table \ref{tab:deployability}.
+We found that 100\% of tested fixed-line networks
 delivered inconsistent retransmissions successfully. However, the delivery
@@
 loss.
+
+\begin{table}
+  \centering
+    \begin{tabular}{llcc}
+    \toprule
+    & ISP & Port 4001 & Port 80 \\ \midrule
+    \textbf{Fixed-line} & Andrews \& Arnold & \pass & \pass \\
+    & BT & \pass & \pass \\
+    & Demon & \pass & \pass \\
+    & EE & \pass & \pass \\
+    & Eclipse & \pass & \pass \\
+    & Sky & \pass & \pass \\
+    & TalkTalk & \pass & \pass \\
+    & Virgin Media & \pass & \pass \\ \midrule
+    \textbf{Cellular} & EE & \okay & \okay \\
+    & O2 & \okay & \okay \\
+    & Three & \pass & \pass \\
+    & Vodafone & \pass & \okay \\
+    \bottomrule
+    \end{tabular}
+  \caption[]{Deployability of inconsistent retransmissions, where \pass \thinspace
+  indicates successful delivery, \okay \thinspace indicates delivery of the original
+  data, and \fail \thinspace indicates connection failure (none observed)}
+  \label{tab:deployability}
+\end{table}
 

### 2016-05-23 paper.tex

--- 2016-05-23 paper.tex
@@
 \usepackage{tabularx}
+\usepackage{rotating}
+\usepackage{tikz}
+
+\makeatletter
+\global\let\tikz@ensure@dollar@catcode=\relax
+\makeatother
+
+\newcommand{\fail}{\tikz\draw[red,fill=red] (0,0) circle (.5ex);}
+\newcommand{\pass}{\tikz\draw[green,fill=green] (0,0) circle (.5ex);}
+\newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) circle (.5ex);}
+
 \usepackage{xspace}

### 2016-05-23 paper.tex
2.3

--- 2016-05-23 paper.tex
@@
 \subsection{Message-oriented Dependencies}
-Given deadlines and partial reliability, the next important transport
-service is \emph{dependency management}. There is no point in sending data
-that cannot be used because it depends on previous data that was not
-received. Providing this service is complicated by the two ways in which
-data can be \emph{useful} to applications: it may itself by played out, or
-it may be needed as part of the application's decoding chain. For example,
-an I frame provides utility both in being played out, and in allowing for
-subsequent P and B frames. As a result, dependency information may result
-in deadline information being overruled: data may be sent even if it cannot
-be played out, if it is necessary for the decoding of other data.
 
-Providing partial reliability, through deadlines and dependencies, requires
-the use of application-level framing \cite{clark:1990:architecture} to
-make the best use of payload data, irrespective of packet loss. At the
-transport layer, this implies that we provide a \emph{message oriented}
-service, that maintains ADU boundaries.
+The combination of deadlines and partial reliability leads to \emph{dependency
+management} as an important transport service.
+% There is no point in sending data that cannot be used because it depends on
+% previous data that was not received.
+In particular, data should never be sent when it relies on a previous
+transmission that was never received. Providing this service is complicated by
+the two ways in which data can be \emph{useful} to applications: It may itself
+by played out, or it may be needed as part of the application's decoding chain.
+For example, I frames provide utility both in being played out, and as
+foundations for subsequent P and B frames. As a result, dependencies may take
+higher priority than deadlines.
+% : data may be sent even if it
+% cannot be played out, if it is necessary for the decoding of other data.
 
-Many multimedia applications make use of multiple data flows. For example,
-a simple IPTV application will maintain separate audio and video flows.
-These could be sent across multiple transport-layer connections, but
-overheads can be reduced by multiplexing these flows on a single
-connection. A \emph{sub-stream} service allows this.
+Partial reliability in the context of both deadlines and dependencies
+requires application-level framing \cite{clark:1990:architecture} to
+make the best use of payload data, irrespective of packet loss. At the transport
+layer, this implies that we provide a \emph{message oriented} service, that
+maintains ADU \todo{first mention?} boundaries.
+
+Message orientation may also be used to construct a \emph{sub-stream} service.
+Many multimedia applications make use of multiple data streams. For example, a
+simple IPTV application will maintain separate audio and video stream. These
+could be sent across multiple transport-layer connections, but overheads can be
+reduced by multiplexing these flows on a single connection.
 
@@
 
-The architecture of TCP Hollywood is shown in Figure
-\ref{diagram:tcp-hollywood}. TCP Hollywood implements all of the services
-described in Section \ref{sec:services}, splitting functionality across
-an intermediary layer in user-space, and a set of modifications to the
-kernel. This split allows applications to program against one API, whether
-or not the kernel modifications are available: the intermediary layer
-functions in both cases.
+The architecture of TCP Hollywood is shown in
+Figure \ref{diagram:tcp-hollywood}. TCP Hollywood provides all of the
+services described in Section \ref{sec:services}, with functionality split
+across an intermediary layer (in user-space), and a set of modifications to
+the kernel. The intermediary layer is responsible for providing the
+message-oriented abstraction, and exposing the API for partial reliability.
+The kernel modifications implement the deadline and dependency services,
+as well as enabling out-of-order delivery of incoming segments. This split
+allows applications to use a single API, whether or not the kernel
+modifications are available. TCP Hollywood inherits standard TCP's
+congestion control mechanism. Clearly, this loss-based mechanism is not
+suitable for real-time applications, and a delay-based algorithm should
+be adopted.
 
-At the sender, applications pass messages (using an API similar to that
-given in Table \ref{tab:api}) to the intermediary layer, with their
-metadata, including deadline and dependency information. At the 
-intermediary layer, COBS encoding \cite{CB97COBS} is used to escape all
-zero bytes in the message data, allowing them to be used as framing
-markers. The message's metadata is then attached to the encoded and framed
-message, before being passed to the kernel using the standard Berkeley
-sockets API.
+\todo{Stephen: give more detail, and expand this to a whole column}
 
-At the kernel, the message data is queued in TCP's sending buffer, while
-the metadata is held in a separate structure. Nagle's algorithm, designed
-to coalesce smaller writes into larger segments, is disabled to minimise
-latency. As segments are (re-)transmitted, their deadlines and dependencies
-are checked to ensure that the message will be useful on arrival. In the
-current version of TCP Hollywood, the dependency check does not overrule
-the deadline check: only data that can be played out will be sent. If the
-message does not pass the liveness check, the next message in the queue
-that is live will be sent instead. If this is a retransmission, then
-inconsistent retransmissions will be used: the replacement message will be
-sent with the same TCP sequence number as the original.
-
-At the receiver, segments are passed to the kernel, where they are
-initially processed as under standard TCP: duplicate acknowledgements are
-generated for out-of-order segments, for example. After this, a metadata
-entry is created, and placed in FIFO queue. When the intermediary layer
-reads from the socket, it receives the segment associated with the
-metadata entry at the head of the queue, with its TCP sequence number
-attached.
-
-At the intermediary layer on the receiver, incoming segments are scanned
-for complete messages (i.e., data between two zero bytes), which are
-decoded and passed to the application. Any message fragments are buffered,
-alongside their TCP sequence number, awaiting the arrival of the remainder
-of the message. Once the message has been reassembled, it is decoded, and
-delivered to the application.
-
-TCP Hollywood inherits standard TCP's congestion control mechanism.
-Clearly, being a loss-based algorithm, this is not appropriate for
-real-time applications.
 \todo{discuss how we need to change TCP congestion control}

### 2016-05-23 paper.tex
2.4

@@
 
+Table \ref{tab:services} summarises the transport services discussed in this
+Section, and required for real-time multimedia applications.
 %--------------------------------------------------------------------------------------------------
@@
 \subsection{Connection and Congestion Control}
-We note that \emph{congestion control} is important. Clearly, the historic
-belief that real-time applications require an isochronous channel, and
-cannot be subjected to congestion control, is not practical on the
-Internet. Further, while there is a number of applications that are not
-adaptive (and require constant bandwidth), an increasing number are
-adaptive, and would better serve users by adapting to the available
-bandwidth.
 
-We note that a connection-oriented transport is not a core requirement for
-many real-time multimedia applications.  Indeed, flexibility to change the
-destination of data within a call could be beneficial for applications that
-support mobile users, or for some forms of multiparty session. However, to
-support NAT traversal and to help dynamically manage firewall pinholes, it
-is often desirable for the transport to be connection oriented. We find
-these concerns outweigh the benefits of connectionless transport, and so
-add a requirement for connection oriented service.
-Similarly, while not strictly needed by the applications, it is beneficial
-if the transport provides a keep-alive service to refresh NAT and firewall
-bindings if the application goes silent.
+We note the importance of congestion control. %Clearly,
+The historic belief that real-time applications require an isochronous channel,
+and should not be subjected to congestion control, is impractical on the
+Internet. Further, while some applications are non-adaptive or constant bitrate,
+an increasing number are either or both of adaptive and variable bitrate. Such
+applications would better serve users by adapting to the available bandwidth.
 
-Table \ref{tab:services} summarises the transport services required for
-real-time multimedia applications.
+We note that a connection-oriented transport is a lesser, but useful,
+requirement.
+% for many real-time multimedia applications.
+Indeed, flexibility to
+change the destination of data within a call could be beneficial for
+applications that support mobile users, or for some forms of multiparty session.
+However, to support NAT traversal and to help dynamically manage firewall
+pinholes, it is often desirable for the transport to be connection oriented. We
+believe these concerns outweigh the benefits of connectionless transport, and so
+add a requirement for connection oriented service. Similarly, while not strictly
+needed by the applications, it is beneficial if the transport provides a
+keep-alive service to refresh NAT and firewall bindings if the application goes
+silent.
 

### 2016-05-23 paper.tex
table font size
one more font size
Work on Conclusions

--- 2016-05-23 paper.tex
@@
 
-\newcommand{\ie}{{i.e.}\xspace}
+\newcommand{\ie}{{i.e.,}\xspace}
 \newcommand*\rot{\rotatebox{90}}
@@
 
-to make effective use of the network, need to deploy new transport services
-and protocols
+The standard transport protocols, TCP and UDP, are not well-suited for
+real-time applications. Both can be made to work, but the existence of
+numerous papers exploring how to make media play-out over TCP reliable,
+and almost as extensive a collection discussing UDP-based protocol design,
+suggests that this is difficult to do well. To make effective use of the
+network, simplify real-time application design and implementation, so it
+doesn't require expert transport protocol knowledge to build a new
+application, and hence support innovation, we need to deploy new transport
+services and protocols.
 

### 2016-05-23 paper.tex

--- 2016-05-23 paper.tex
@@
 \begin{table}
+  \small
   \centering
@@
 \begin{table*}
+  \small
   \centering
@@
 \begin{table}
-  \centering
+  \small \centering
     \begin{tabular}{llcc}

### 2016-05-23 paper.tex
Resolves ADU to-do
Conclusions

--- 2016-05-23 paper.tex
@@
 suggests that this is difficult to do well. To make effective use of the
-network, simplify real-time application design and implementation, so it
-doesn't require expert transport protocol knowledge to build a new
-application, and hence support innovation, we need to deploy new transport
-services and protocols.
+network, and simplify real-time application design and implementation, we
+need to deploy new transport services and protocols that allow innovative
+applications to be developed by users who are not experts in transport 
+protocol design. We discussed requirements such for a new transport, in
+the context of the TAPS framework, and outlined a straw-man abstract API,
+in Section \ref{sec:services}.
 
-seems likely that the appropriate long-term approach for doing this is to
-repurpose UDP as a demultiplexing layer for higher-layer protocols.
+It seems likely that the right long-term approach for doing this is to
+repurpose UDP as a demultiplexing layer for higher-layer protocols. We can
+then deploy an appropriate transport protocol framework as a user-space
+library, that can be reused as appropriate. In the short term, however, 
+there are sufficient networks that block UDP, that any new transport
+protocol needs to be able to run over TCP. Sections \ref{sec:ossification}
+and \ref{sec:partial} discuss how this can be done, and suggest from some
+initial measurement studies that this may be feasible to deploy. Section
+\ref{sec:realising} considers prototypes that present such services over
+UDP, and presents our initial prototype demonstrated for TCP-based use.
 
-in the short term, we need to deal with networks that block UDP, and be
-able to run over TCP.
-our initial work shows that this is frequently possible, and works well
-enough to offer benefit.
+Making effective use of the TAPS concept to support real-time multimedia
+requires making significant changes to the TCP specification and API, if 
+such applications are to be able to fall-back to TCP from UDP, and still
+achieve their performance needs.
 
@@
 %\section{Acknowledgements}
-
+%
 % Acknowledge funding sources.
-
+%
 %==================================================================================================

### 2016-05-23 paper.tex

--- 2016-05-23 paper.tex
@@
 layer, this implies that we provide a \emph{message oriented} service, that
-maintains ADU \todo{first mention?} boundaries.
+maintains application data unit (ADU) boundaries.
 
@@
 
-to make effective use of the network, need to deploy new transport services
-and protocols
+The standard transport protocols, TCP and UDP, are not well-suited for
+real-time applications. Both can be made to work, but the existence of
+numerous papers exploring how to make media play-out over TCP reliable, 
+and almost as extensive a collection discussing UDP-based protocol design,
+suggests that this is difficult to do well. To make effective use of the
+network, and simplify real-time application design and implementation, we
+need to deploy new transport services and protocols that allow innovative
+applications to be developed by users who are not experts in transport 
+protocol design. We discussed requirements such for a new transport, in
+the context of the TAPS framework, and outlined a straw-man abstract API,
+in Section \ref{sec:services}.
 
-seems likely that the appropriate long-term approach for doing this is to
-repurpose UDP as a demultiplexing layer for higher-layer protocols.
+It seems likely that the right long-term approach for doing this is to
+repurpose UDP as a demultiplexing layer for higher-layer protocols. We can
+then deploy an appropriate transport protocol framework as a user-space
+library, that can be reused as appropriate. In the short term, however, 
+there are sufficient networks that block UDP, that any new transport
+protocol needs to be able to run over TCP. Sections \ref{sec:ossification}
+and \ref{sec:partial} discuss how this can be done, and suggest from some
+initial measurement studies that this may be feasible to deploy. Section
+\ref{sec:realising} considers prototypes that present such services over
+UDP, and presents our initial prototype demonstrated for TCP-based use.
 
-in the short term, we need to deal with networks that block UDP, and be
-able to run over TCP.
-our initial work shows that this is frequently possible, and works well
-enough to offer benefit.
+Making effective use of the TAPS concept to support real-time multimedia
+requires making significant changes to the TCP specification and API, if 
+such applications are to be able to fall-back to TCP from UDP, and still
+achieve their performance needs.
 

### 2016-05-23 paper.tex
Tweak metadata
Space saving

@@
 
-\ccsdesc[500]{Networks~Network protocol design}
+\ccsdesc[500]{Networks~Protocol design}
 \ccsdesc[500]{Networks~Transport protocols}
@@
 
-    \todo{relate this to the transport services described earlier}
+%    \todo{relate this to the transport services described earlier}
 
@@
   \end{tabular}
-  \caption{Transport services required for real-time multimedia}
+  \caption{Transport services for real-time multimedia}
   \label{tab:services}
@@
 API, which provides \textit{timed reliability}, using a ``lifetime''
-specified by the application. \todo{CSP: check this}
+specified by the application. 
+%\todo{CSP: check this}
 
@@
 These deployment experiments suggest that our protocol should be
-flexible: inconsistent retransmissions may not be delivered, and we should
+flexible: inconsistent retransmissions might not be delivered, and we should
 handle reception of the original segment. If the protocol detects that
-inconsistent retransmissions are not being delivered, then the mechanism
+inconsistent retransmissions are not being delivered, they
 can be disabled for the connection. Further, if a connection reset occurs,
@@
 
-The use of inconsistent retransmissions may interact negatively with
-caching and resegmenting middleboxes, resulting in the corruption of
+Use of inconsistent retransmissions may interact negatively with
+caching and re-segmenting middleboxes, resulting in the corruption of
 messages between sender and receiver. A message may be formed from the
@@
 integrity. The role of a checksum may also be fulfilled by using a secure
-transport protocol, such as DTLS \cite{rfc:6347}.
+transport, such as DTLS \cite{rfc:6347}.
 
@@
 \cite{RFC4566} offer/answer exchange \cite{RFC3264} as part of a WebRTC
-session \cite{jennings:2013:rtcweb} (the WebRTC media channel uses RTP
-running over UDP also, further showing the utility of UDP-based data).
+session \cite{jennings:2013:rtcweb} (WebRTC media uses RTP
+over UDP also, further showing the utility of UDP-based data).
 This has been deployed in popular web browsers, with global deployment, and
@@
  \includegraphics[scale=0.45]{figures/tcp-hollywood.pdf}
- \caption{TCP Hollywood architecture \todo{too similar to NOSSDAV fig 1?}}
+  \caption{TCP Hollywood architecture}
 \label{diagram:tcp-hollywood}

### 2016-05-23 paper.tex
Related work

@@
 
-The Minion protocol suite \cite{nowlan:2012:minion} includes uTCP, a
-protocol that uses TCP as a substrate to provide an unordered,
-message-oriented service to applications. While this enables many of
-the transport services described in Section \ref{sec:services}, support for
-partial reliability, deadlines, and dependencies is not provided.
+Related changes to TCP are made by Minion protocol \cite{nowlan:2012:minion},
+that uses TCP as a substrate to provide an unordered, message-oriented
+service to applications, enabling some of the transport services described
+in Section \ref{sec:services}, but without support for partial reliability,
+deadlines, and dependencies.
 Time-Lined TCP (TLTCP) \cite{mukherjee:2000:timelines} similarly provides a
@@
 their time-line, after which they are discarded. The mechanism by which
-this service is provided (i.e., the introduction of gaps in the sequence
-space) reduces deployability.
+this service is provided (introducing gaps in the sequence space) hinders
+deployment.
 
-\todo{QUIC}
-
-\todo{SPUD}
-
-\todo{WebRTC data channel - SCTP over UDP}
-
-\todo{RFC6773 - DCCP over UDP}
+QUIC \cite{draft-tsvwg-quic-protocol-02} demonstrates that similar services
+can be provided by a new protocol running over UDP, while \cite{RFC6773} and
+\cite{draft-ietf-rtcweb-data-channel-13} demonstrate that existing protocols,
+DCCP and SCTP, can also be effectively tunnelled over UDP. Fallback to TCP
+is discussed in this paper, and on our previous work \cite{mcquistin2016hollywood}.
 

### 2016-05-23 paper.tex
Move some text from Realising to Conclusions.

@@
 
-TCP Hollywood inherits standard TCP's congestion control mechanism.
-Clearly, being a loss-based algorithm, this is not appropriate for
-real-time applications.
-\todo{discuss how we need to change TCP congestion control}
+% TCP Hollywood inherits standard TCP's congestion control mechanism.
+% Clearly, being a loss-based algorithm, this is not appropriate for
+% real-time applications.
+% \todo{discuss how we need to change TCP congestion control}
 
@@
 possible.
-
-The challenge for the future is in combining such techniques below a common
-API, so that an application can transparently switch between UDP-based and
-TCP-based transport, depending on what is supported by the underlying
-network. This is the promise of the TAPS API, that we have shown ought to
-be feasible for real-time applications.
 
@@
 
-Making effective use of the TAPS concept to support real-time multimedia
-requires making significant changes to the TCP specification and API, if 
-such applications are to be able to fall-back to TCP from UDP, and still
-achieve their performance needs.
+The challenge for the future is in combining such techniques below a common
+API, so that an application can transparently switch between UDP-based and
+TCP-based transport, depending on what is supported by the underlying
+network. This is the promise of the TAPS API, that we have shown ought to
+be feasible for real-time applications.
+
+% Making effective use of the TAPS concept to support real-time multimedia
+% requires making significant changes to the TCP specification and API, if 
+% such applications are to be able to fall-back to TCP from UDP, and still
+% achieve their performance needs.
 

### 2016-05-23 paper.tex
Space saving

@@
 implement alternatives that seek stability, or compatibility with the
-dictates of a video codec, rather than the traditional ``TCP Friendly''
+dictates of a video codec, rather than traditional ``TCP Friendly''
 congestion control -- even if implemented within TCP.
@@
 
-These changes could address many of the transport service needs we have for
+These changes could address many of the transport service needs for
 real-time applications, but still leave a critical issue of how to improve
@@
 
-When a TCP segment is to be retransmitted, the current mapping between its
+When a TCP segment is to be retransmitted, the mapping between its
 sequence number and application-level sequence numbers is used to determine

### 2016-05-23 paper.tex
todos and space saving

@@
 \usepackage{tikz}
+\usepackage{xspace}   %MF
+\usepackage{enumitem} %MF
 
@@
 \newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) circle (.5ex);}
-
-\usepackage{xspace}
 
@@
 %to show that, despite network ossification,
-the network has the flexibility to deploy new transport protocols,
-provided care is taken to reinterpret layer boundaries that respect
-conventional UDP and TCP.\todo{I can infer the meaning of previous sentence having worked on the problem; otherwise may be very ambiguous to the reader.}
+the network has the flexibility to deploy new transport protocols, provided care
+is taken to reinterpret application and transport layer boundaries that respect
+conventional UDP and TCP.
 %and deploy those new transports in the context of TCP and UDP.
@@
 
-Networked multimedia deadlines are unusual when compared to other real-time
-systems. They are simultaneously flexible and strict: flexible in that the exact
-value of the deadline is typically not important, provided it is of the right
-order-of-magnitude for the application, but strict in that any particular
-deadline provides a hard cut-off, after which the data is useless.
-\todo{Is this para necessary at all?
-Its contribution is unclear and/or unconvincing -- flexible is fine; strict is less believable, \ie
-how are RT systems any different when the deadline expires? (Many RTSs will
-discard, too. Ultimately, I get what it is trying to say; I'm less sure it needs saying) Maybe the following(?):}
-% When compared to conventional real-time systems, networked multimedia
-% deadlines are neither soft nor hard.
-It is useful to draw a distinction between in networked multimedia deadlines and
-deadlines in conventional real-time systems.
-A multimedia deadline is neither soft, since arrivals played after a deadline contribute to poor experience; nor are deadlines hard, since occasional missed messages can be tolerated.
+% % When compared to conventional real-time systems, networked multimedia
+% % deadlines are neither soft nor hard.
+% Networked multimedia deadlines are unusual when compared to other real-time
+% systems. They are simultaneously flexible and strict: flexible in that the exact
+% value of the deadline is typically not important, provided it is of the right
+% order-of-magnitude for the application, but strict in that any particular
+% deadline provides a hard cut-off, after which the data is useless.
+It is useful to distinguish between networked multimedia deadlines and
+deadlines in conventional real-time systems. A multimedia deadline is neither
+soft, since arrivals played after a deadline contribute to poor experience; nor
+are deadlines hard, since occasional missed messages can be tolerated.
 
 \subsection{Partial Reliability}
-In a network that offers only a best-effort packet delivery service, deadlines
-constrain delivery to \emph{partial reliability}.
+In a best-effort network, deadlines
+constrain packet delivery service to \emph{partial reliability}.
 %The network has non-zero probability of losing any particular packet.
@@
 
-Partial reliability in the context of both deadlines and dependencies
+In the context of both deadlines and dependencies coupled with packet loss, partial reliability
 requires application-level framing \cite{clark:1990:architecture} to
-make the best use of payload data, irrespective of packet loss. At the transport
-layer, this implies that we provide a \emph{message oriented} service, that
+make the best use of payload data. At the transport
+layer, this implies a \emph{message oriented} service, that
 maintains application data unit (ADU) boundaries.
@@
 Indeed, flexibility to
-change the destination of data within a call could be beneficial for
+change the destination within a call is beneficial for
 applications that support mobile users, or for some forms of multiparty session.
@@
 
-\begin{itemize}
+\begin{itemize}[nosep]
   \item Hosts setup and tear-down sockets using the \texttt{socket()}
@@
 
-    The \texttt{send\_message()} call takes four additional parameters
-    compared to the standard Berkeley Sockets API. These are 1) a
-    message sequence number, that can be used to re-order messages and
-    detect message loss; 2) a relative deadline, which is combined
-    with an estimate of the current round-trip-time, and the time that
-    the message has spent in the sending buffer, to determine if a
-    message will arrive in time to be played-out; 3) the message
-    sequence number of any message on which this depends, for example,
-    of a video I-frame on which a P-frame is predicted; and 4) a
-    sub-stream identifier, used, for example, to differentiate audio,
-    video, sub-title, control, and repair streams.
+\end{itemize}
 
-    The \texttt{recv\_message()} call returns the sequence number,
-    dependency information, and sub-stream identifier along with any
-    received message, allowing the receiver to direct it to the correct
-    decoding queue.
+    It is instructive to compare the partially reliable send and receive
+    functions to their TCP-like counterparts. The \texttt{send\_message()} call
+    takes four additional parameters compared to the standard Berkeley Sockets
+    API. These are 1) a message sequence number, that can be used to re-order
+    messages and detect message loss; 2) a relative deadline, which is combined
+    with an estimate of the current round-trip-time, and the time that the
+    message has spent in the sending buffer, to determine if a message will
+    arrive in time to be played-out; 3) the message sequence number of any
+    message on which this depends, for example, of a video I-frame on which a
+    P-frame is predicted; and 4) a sub-stream identifier, used, for example, to
+    differentiate audio, video, sub-title, control, and repair streams.
+
+    The \texttt{recv\_message()} call returns the sequence number, dependency
+    information, and sub-stream identifier along with any received message,
+    allowing the receiver to direct it to the correct decoding queue.
 
 %    \todo{relate this to the transport services described earlier}
-
-\end{itemize}
 

### 2016-05-23 paper.tex
last todos resolved (commented out; saw no need to address them)

@@
 be exchanged.
-\todo{expand, and discuss how this evolution has occurred to date}
+%\todo{expand, and discuss how this evolution has occurred to date}
 We note that, while standardised TCP congestion control has followed the

### 2016-05-23 paper.tex
small font for refs
\balance
small space saving
Fix borked merge
more space

@@
 We note the importance of congestion control. %Clearly,
-The historic belief that real-time applications require an isochronous channel,
-and should not be subjected to congestion control, is impractical on the
+Historically real-time applications needed an isochronous channel,
+and needed to avoid subjected to congestion control. This is impractical on the
 Internet. Further, while some applications are non-adaptive or constant bitrate,
-an increasing number are either or both of adaptive and variable bitrate. Such
-applications would better serve users by adapting to the available bandwidth.
+an increasing number are either or both of adaptive and variable bitrate. Users would be better served by
+applications that to available bandwidth.
 
@@
 
-   \item Once the connection is established, the receiver application
-     indicates the media play-out delay it is using, in milliseconds, via
+   \item Once the connection is established, the receiver
+     indicates the its media play-out delay, in milliseconds, via
      the \texttt{set\_po\_delay()} call. This specifies the time that the
@@
 \begin{table}
-  \small
+  \scriptsize
   \centering
@@
 \begin{table}
-  \small \centering
+  \scriptsize \centering
     \begin{tabular}{llcc}
@@
 %==================================================================================================
+\balance
 \bibliographystyle{abbrv}

### 2016-05-23 paper.tex
space

@@
     It is instructive to compare the partially reliable send and receive
-    functions to their TCP-like counterparts. The \texttt{send\_message()} call
-    takes four additional parameters compared to the standard Berkeley Sockets
-    API. These are 1) a message sequence number, that can be used to re-order
-    messages and detect message loss; 2) a relative deadline, which is combined
-    with an estimate of the current round-trip-time, and the time that the
-    message has spent in the sending buffer, to determine if a message will
-    arrive in time to be played-out; 3) the message sequence number of any
-    message on which this depends, for example, of a video I-frame on which a
-    P-frame is predicted; and 4) a sub-stream identifier, used, for example, to
-    differentiate audio, video, sub-title, control, and repair streams.
+    functions to their Berkeley Sockets API counterparts. The
+    \texttt{send\_message()} call takes four additional parameters. These are 1)
+    a message sequence number, that can be used to re-order messages and detect
+    message loss; 2) a relative deadline, which is combined with an estimate of
+    the current round-trip-time, and the time that the message has spent in the
+    sending buffer, to determine if a message will arrive in time to be
+    played-out; 3) the message sequence number of any message on which this
+    depends, for example, of a video I-frame on which a P-frame is predicted;
+    and 4) a sub-stream identifier, used, for example, to differentiate audio,
+    video, sub-title, control, and repair streams.
 

### 2016-05-23 paper.tex
Fixes typos

--- 2016-05-23 paper.tex
@@
 \usepackage{tikz}
-\usepackage{xspace}   %MF
-\usepackage{enumitem} %MF
 
@@
 \newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) circle (.5ex);}
+
+\usepackage{xspace}
 
@@
 provided by widely-deployed transport protocols. Ossification prevents the
-deployment of a novel protocols, restricting solutions to protocols using
+deployment of novel protocols, restricting solutions to protocols using
 either TCP or UDP as a substrate. We describe the transport services
-required by real-time applications. We show that, in the short term (i.e.,
+required by real-time applications. We show that, in the short-term (i.e.,
 while UDP is blocked at current levels), TCP offers a feasible substrate
@@
 %to show that, despite network ossification,
-the network has the flexibility to deploy new transport protocols, provided care
-is taken to reinterpret application and transport layer boundaries that respect
-conventional UDP and TCP.
+the network has the flexibility to deploy new transport protocols,
+provided care is taken to reinterpret layer boundaries that respect
+conventional UDP and TCP.\todo{I can infer the meaning of previous sentence having worked on the problem; otherwise may be very ambiguous to the reader.}
 %and deploy those new transports in the context of TCP and UDP.
@@
 
-% % When compared to conventional real-time systems, networked multimedia
-% % deadlines are neither soft nor hard.
-% Networked multimedia deadlines are unusual when compared to other real-time
-% systems. They are simultaneously flexible and strict: flexible in that the exact
-% value of the deadline is typically not important, provided it is of the right
-% order-of-magnitude for the application, but strict in that any particular
-% deadline provides a hard cut-off, after which the data is useless.
-It is useful to distinguish between networked multimedia deadlines and
-deadlines in conventional real-time systems. A multimedia deadline is neither
-soft, since arrivals played after a deadline contribute to poor experience; nor
-are deadlines hard, since occasional missed messages can be tolerated.
+Networked multimedia deadlines are unusual when compared to other real-time
+systems. They are simultaneously flexible and strict: flexible in that the exact
+value of the deadline is typically not important, provided it is of the right
+order-of-magnitude for the application, but strict in that any particular
+deadline provides a hard cut-off, after which the data is useless.
+\todo{Is this para necessary at all?
+Its contribution is unclear and/or unconvincing -- flexible is fine; strict is less believable, \ie
+how are RT systems any different when the deadline expires? (Many RTSs will
+discard, too. Ultimately, I get what it is trying to say; I'm less sure it needs saying) Maybe the following(?):}
+% When compared to conventional real-time systems, networked multimedia
+% deadlines are neither soft nor hard.
+It is useful to draw a distinction between in networked multimedia deadlines and
+deadlines in conventional real-time systems.
+A multimedia deadline is neither soft, since arrivals played after a deadline contribute to poor experience; nor are deadlines hard, since occasional missed messages can be tolerated.
 
 \subsection{Partial Reliability}
-In a best-effort network, deadlines
-constrain packet delivery service to \emph{partial reliability}.
+In a network that offers only a best-effort packet delivery service, deadlines
+constrain delivery to \emph{partial reliability}.
 %The network has non-zero probability of losing any particular packet.
@@
 
-In the context of both deadlines and dependencies coupled with packet loss, partial reliability
+Partial reliability in the context of both deadlines and dependencies
 requires application-level framing \cite{clark:1990:architecture} to
-make the best use of payload data. At the transport
-layer, this implies a \emph{message oriented} service, that
+make the best use of payload data, irrespective of packet loss. At the transport
+layer, this implies that we provide a \emph{message oriented} service, that
 maintains application data unit (ADU) boundaries.
@@
 Indeed, flexibility to
-change the destination within a call is beneficial for
+change the destination of data within a call could be beneficial for
 applications that support mobile users, or for some forms of multiparty session.
@@
 
-\begin{itemize}[nosep]
+\begin{itemize}
   \item Hosts setup and tear-down sockets using the \texttt{socket()}
@@
 
-\end{itemize}
+    The \texttt{send\_message()} call takes four additional parameters
+    compared to the standard Berkeley Sockets API. These are 1) a
+    message sequence number, that can be used to re-order messages and
+    detect message loss; 2) a relative deadline, which is combined
+    with an estimate of the current round-trip-time, and the time that
+    the message has spent in the sending buffer, to determine if a
+    message will arrive in time to be played-out; 3) the message
+    sequence number of any message on which this depends, for example,
+    of a video I-frame on which a P-frame is predicted; and 4) a
+    sub-stream identifier, used, for example, to differentiate audio,
+    video, sub-title, control, and repair streams.
 
-    It is instructive to compare the partially reliable send and receive
-    functions to their TCP-like counterparts. The \texttt{send\_message()} call
-    takes four additional parameters compared to the standard Berkeley Sockets
-    API. These are 1) a message sequence number, that can be used to re-order
-    messages and detect message loss; 2) a relative deadline, which is combined
-    with an estimate of the current round-trip-time, and the time that the
-    message has spent in the sending buffer, to determine if a message will
-    arrive in time to be played-out; 3) the message sequence number of any
-    message on which this depends, for example, of a video I-frame on which a
-    P-frame is predicted; and 4) a sub-stream identifier, used, for example, to
-    differentiate audio, video, sub-title, control, and repair streams.
-
-    The \texttt{recv\_message()} call returns the sequence number, dependency
-    information, and sub-stream identifier along with any received message,
-    allowing the receiver to direct it to the correct decoding queue.
+    The \texttt{recv\_message()} call returns the sequence number,
+    dependency information, and sub-stream identifier along with any
+    received message, allowing the receiver to direct it to the correct
+    decoding queue.
 
 %    \todo{relate this to the transport services described earlier}
+
+\end{itemize}
 
@@
     \toprule
-      Transport Service            & Function & Parameters                                     & Return Value \\
+      Transport Service            & Function & Parameters                                     & Return Value(s) \\
     \midrule
@@
 be exchanged.
-%\todo{expand, and discuss how this evolution has occurred to date}
+\todo{expand, and discuss how this evolution has occurred to date}
 We note that, while standardised TCP congestion control has followed the
@@
 applications to be developed by users who are not experts in transport 
-protocol design. We discussed requirements such for a new transport, in
+protocol design. We discussed requirements for such a new transport, in
 the context of the TAPS framework, and outlined a straw-man abstract API,
@@
 then deploy an appropriate transport protocol framework as a user-space
-library, that can be reused as appropriate. In the short term, however,
+library, that can be reused as appropriate. In the short-term, however, 
 there are sufficient networks that block UDP, that any new transport

### 2016-05-23 paper.tex
space

--- 2016-05-23 paper.tex
@@
 
-The API that is exposed to applications using TCP is not visible to the
-network, and can be changed. Relaxing the API to enable the out-of-order
+The API that is exposed to applications using TCP is invisible from the
+network, and can be changed. Relaxing the API to enable out-of-order
 delivery of segments is trivial: segments are delivered to the application

### 2016-06-21 paper.tex
Adds back some space, to shift Section 4 title
Add ACM bibstrip
Adds to-dos based on reviews

@@
 
-\toappear{}
+\CopyrightYear{2016} 
+\setcopyright{acmlicensed}
+\conferenceinfo{ANRW '16,}{July 16 2016, Berlin, Germany}
+\isbn{978-1-4503-4443-2/16/07}\acmPrice{\$15.00}
+\doi{http://dx.doi.org/10.1145/2959424.2959443}
+
 \maketitle
@@
 transmission that was never received. Providing this service is complicated by
-the two ways in which data can be \emph{useful} to applications: It may itself
-by played out, or it may be needed as part of the application's decoding chain.
+the two ways in which data can be \emph{useful} to applications: it may itself
+be played out, or it may be needed as part of the application's decoding chain.
 For example, I frames provide utility both in being played out, and as
@@
 higher priority than deadlines.
+\todo{Review A: Some video-specific jargon (I, P, B frames) could be
+expanded/explained.}
 % : data may be sent even if it
@@
 Internet. Further, while some applications are non-adaptive or constant bitrate,
-an increasing number are either or both of adaptive and variable bitrate. Users would be better served by
-applications that to available bandwidth.
+an increasing number are either or both of adaptive and variable bitrate.
+Users would be better served by applications that adapt to available bandwidth.
 
@@
 silent.
+\todo{Review A: Indicate that connectionless protocols can create in-network
+state.}
 
@@
    \item Once the connection is established, the receiver
-     indicates the its media play-out delay, in milliseconds, via
+     then indicates its media play-out delay, in milliseconds, via
      the \texttt{set\_po\_delay()} call. This specifies the time that the
@@
 congestion control -- even if implemented within TCP.
+\todo{Review A: Show how the API supports alternative congestion control
+algorithms.}
+\todo{Review B: Alternatives might have deployment challenges (e.g.,
+fairness between delay and loss based algorithms).}
 
@@
 dependencies.
+
+\todo{Review A: Section 3 should also note that connectionless send and
+receive already work on connection-oriented sockets}
 
@@
 required to allow messages to be uniquely identified.
+\todo{Review B: Inconsistent retransmissions are not well introduced. Should
+be more explicit, and earlier.}
 
@@
 
+\todo{Review A: Section 4 should have a forward reference to section 5}
 %==================================================================================================
@@
 delivered to the application.
-
+\todo{Review B: Head-of-line blocking problem not introduced, so not clear
+why this mechanism is needed.}
 % TCP Hollywood inherits standard TCP's congestion control mechanism.
@@
 DCCP and SCTP, can also be effectively tunnelled over UDP. Fallback to TCP
-is discussed in this paper, and on our previous work \cite{mcquistin2016hollywood}.
+is discussed in this paper, and on our previous work \cite{mcquistin2016tcp}.
 

### 2016-06-22 paper.tex
Fixes to-do (expands on I, P, and B frame example)

@@
 be played out, or it may be needed as part of the application's decoding chain.
-For example, I frames provide utility both in being played out, and as
-foundations for subsequent P and B frames. As a result, dependencies may take
-higher priority than deadlines.
-\todo{Review A: Some video-specific jargon (I, P, B frames) could be
-expanded/explained.}
-% : data may be sent even if it
-% cannot be played out, if it is necessary for the decoding of other data.
+Interdependencies between frames of video exist within a number of codecs,
+including MPEG-1 \cite{le1991mpeg}, and newer codecs such as H.264
+\cite{wiegand2003overview}. As an example, we consider MPEG-1 video:
+I-frames are independent, while P- and B-frames contain only the changes 
+since the previous frame (P) or between frames (B), and so are dependent on
+the successful arrival of other frames. An I-frame may not arrive in time
+to be played out, but may need to be sent to ensure that dependent P- and
+B-frames are useful.
 

### 2016-06-22 paper.tex
Fixes to-do

@@
 However, to support NAT traversal and to help dynamically manage firewall
-pinholes, it is often desirable for the transport to be connection oriented. We
+pinholes, it is often desirable, though not necessary, for the transport to be connection oriented. We
 believe these concerns outweigh the benefits of connectionless transport, and so
@@
 silent.
-\todo{Review A: Indicate that connectionless protocols can create in-network
-state.}
 

### 2016-06-22 paper.tex
Switch to the 9pt version of the ACM template
Fix fonts in figure
Add get/setsockopt to API table and accompanying text

@@
 we sketch an abstract API in  Table \ref{tab:api}. The primitives divide
-into four categories:
+into five categories:
 
@@
     API.
+
+  \item Socket options can be set and read using the \texttt{setsockopt()}
+    and \texttt{getsockopt()} functions respectively, again, mirroring the
+    standard Berkeley sockets API. A socket option may be used to select
+    the desired congestion control algorithm (e.g., as with the
+    \texttt{DCCP\_SOCKOPT\_CCID} socket option in DCCP \cite{rfc:4340}).
 
@@
                                               & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+    \midrule
+                                   & \texttt{getsockopt/setsockopt}                                    
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   &          & \texttt{level} -- Protocol level & \\
+                                   &          & \texttt{option} -- Option name & \\
+                                   &          & \texttt{value} -- Option value & \\
+                                   &          & \texttt{len} -- Option length & \\
     \midrule
@@
 congestion control -- even if implemented within TCP.
-\todo{Review A: Show how the API supports alternative congestion control
-algorithms.}
 \todo{Review B: Alternatives might have deployment challenges (e.g.,
@@
 We conducted further deployment experiments using
-inconsistent retransmissions \cite{mcquistin2016hollywood}, testing all
+inconsistent retransmissions \cite{mcquistin2016tcp}, testing all
 major UK providers. The results are shown in Table \ref{tab:deployability}.
@@
 Section \ref{sec:services} in previous work, with our TCP Hollywood
-proposal \cite{mcquistin2016hollywood}.
+proposal \cite{mcquistin2016tcp}.
 
@@
  \centering
- \includegraphics[scale=0.45]{figures/tcp-hollywood.pdf}
+ \includegraphics[scale=0.45]{figures/tcp-hollywood-fixed.pdf}
   \caption{TCP Hollywood architecture}

### 2016-06-22 paper.tex
Restore tables to their normal size
Update introduction

@@
 
-Real-time applications are increasingly
-%important
-present in the Internet.
-%We want to make it easier to write these applications
-Such applications should be easier to program, while also improving
-the quality of experience for users.
- %of those applications
-by lowering latency and increasing robustness.
-%In this, we are
-This limitations of standard Internet transport protocols makes this a
-challenging target. Moreover, the ossified nature of the network makes
-it increasingly difficult to deploy new transports.
+Real-time applications are increasingly present in the Internet. We want
+to make it easier to write these applications, while also improving the
+quality of experience for users by lowering latency and increasing the
+quality and robustness of the media delivery.
+Unfortunately, the limitations of the standard Internet transport protocols
+make this a challenging target, and the ossified nature of the network
+makes it increasingly difficult to deploy new transport protocols.
 
@@
 
-In practice, only UDP and TCP are widely usable in the Internet,
-since remaining protocols are blocked by firewalls and other
+There have been several attempts to standardise and deploy new transport 
+protocols \cite{rfc:4340,RFC4960}.
+In practice, however, only UDP and TCP are widely usable in the Internet,
+since the remaining protocols are blocked by firewalls and other
 middleboxes. UDP exposes the best-effort IP packet delivery service,
-offering the flexibility to develop new protocols. The high costs
-associated involve defining completely new protocol mechanisms. In
-contrast TCP mechanisms are well defined, consisting of sophisticated
+offering the flexibility to develop new protocols, but at the cost of
+requiring new mechanisms to be defined and implemented from scratch.
+In contrast, TCP mechanisms are well defined, consisting of sophisticated
 congestion control coupled with a reliable, ordered, byte stream API.
@@
 to realise real-time services and APIs in the context of both TCP and
-UDP, despite the limitations imposed by their legacies, and by
-middleboxes, and ossification of the network.  Initial experiments
+UDP, despite the limitations imposed by their legacies, by
+middleboxes, and by the ossification of the network.  Initial experiments
 with our implementation suggest that
@@
 the network has the flexibility to deploy new transport protocols, provided care
-is taken to reinterpret application and transport layer boundaries that respect
-conventional UDP and TCP.
+is taken to reinterpret application and transport layer boundaries in a
+manner that is not at odds with conventional UDP and TCP layer boundaries.
 %and deploy those new transports in the context of TCP and UDP.
@@
 
+We structure the remainder of this paper as follows.
 We begin in Section \ref{sec:services} by discussing transport
@@
 \begin{table}
-  \scriptsize
   \centering
@@
 \begin{table}
-  \scriptsize \centering
+  \centering
     \begin{tabular}{llcc}

### 2016-06-22 paper.tex
Use normal font size
Add todo about labels in Table 3
Update formatting of Section 2

@@
 
-Table \ref{tab:services} summarises the transport services discussed in this
-Section, and required for real-time multimedia applications.
 %--------------------------------------------------------------------------------------------------
-\subsection{Timing and Transport Services}
+\subsection{Desirable Transport Services}
 
 The work in TAPS provides a vocabulary for discussing the components of
-transport protocols.  The vocabulary is useful when discussing the needs of
-real-time applications, and the protocols to support them.
+transport protocols. The vocabulary is useful when discussing the needs 
+of real-time applications, and the protocols to support them. In this 
+section, we use this to describe the transport services we believe are
+required for real-time multimedia applications.
+Table \ref{tab:services} summarises the transport services discussed.
 
+\textbf{Timing and Deadlines:}
 Timing is the most salient feature of real-time applications. Since their data
@@
 
-\subsection{Partial Reliability}
+\textbf{Partial Reliability:}
 In a best-effort network, deadlines
@@
 
-\subsection{Message-oriented Dependencies}
-
+\textbf{Message-oriented Dependencies:}
 The combination of deadlines and partial reliability leads to \emph{dependency
@@
 
-\subsection{Connection and Congestion Control}
-
+\textbf{Connection and Congestion Control:}
 We note the importance of congestion control. %Clearly,
@@
 
-\begin{table*}
-  \small
+\begin{table*}[t]
   \centering
@@
   data, and \fail \thinspace indicates connection failure (none observed)}
+  \todo{the labels need to be distinctive shapes, not just different colours}
   \label{tab:deployability}

### 2016-06-22 paper.tex
Address review 2A comment about in-network state

@@
 
-\textbf{Connection and Congestion Control:}
-We note the importance of congestion control. %Clearly,
-Historically real-time applications needed an isochronous channel,
-and needed to avoid subjected to congestion control. This is impractical on the
-Internet. Further, while some applications are non-adaptive or constant bitrate,
-an increasing number are either or both of adaptive and variable bitrate.
-Users would be better served by applications that adapt to available bandwidth.
+\textbf{Connections and Congestion Control:}
+We note the importance of congestion control. Historically, real-time many
+applications have required an isochronous channel, that is not subject to
+congestion control. This is impractical on the Internet. Further, while
+some applications are non-adaptive or constant bit-rate, an increasing
+number are either or both of adaptive and variable bit-rate. Users would
+be better served by applications that adapt to available bandwidth.
 
-We note that a connection-oriented transport is a lesser, but useful,
-requirement.
-% for many real-time multimedia applications.
-Indeed, flexibility to
-change the destination within a call is beneficial for
-applications that support mobile users, or for some forms of multiparty session.
-However, to support NAT traversal and to help dynamically manage firewall
-pinholes, it is often desirable, though not necessary, for the transport to be connection oriented. We
-believe these concerns outweigh the benefits of connectionless transport, and so
+We note that a connection-oriented transport is a lesser requirement for
+many real-time multimedia applications. 
+Indeed, flexibility to change the destination within a call is beneficial
+for applications that support mobile users, and for some forms of multiparty
+session.
+Maintaining per-connection state at the endpoints is helpful for the
+implementation of many forms of congestion control, however. 
+Signalling messages indicating start and end of connections can also ease
+NAT traversal, and help dynamically manage firewall pinholes, by indicating
+when in-network state should be created and can be torn down. 
+Accordingly, it is often desirable for the transport to be connection
+oriented.
+
+We believe these concerns outweigh the benefits of connectionless transport, and so
 add a requirement for connection oriented service. Similarly, while not strictly

### 2016-06-22 paper.tex
Address review 2B point 1, about TCP Vegas

@@
 be exchanged.
-%\todo{expand, and discuss how this evolution has occurred to date}
 We note that, while standardised TCP congestion control has followed the
@@
 is perhaps the best known approach that changes these constraints, with a
-delay-based algorithm that reduces latency. It would also be possible to
+delay-based algorithm that reduces latency, although it is known to be less
+aggressive than standard TCP, and is prone to starvation; FAST TCP 
+\cite{jin:2004:fast-tcp} is a more modern delay-based algorithm that 
+competes well with standard TCP in many environments, and is seeing 
+commercial deployment. 
+It would also be possible to
 implement alternatives that seek stability, or compatibility with the
@@
 congestion control -- even if implemented within TCP.
-\todo{Review B: Alternatives might have deployment challenges (e.g.,
-fairness between delay and loss based algorithms).}
+The development of TCP congestion control shows that there might be
+fairness issues as new algorithms are deployed, but the network does not
+prevent the deployment of those algorithms.
 

### 2016-06-22 paper.tex
Fix todo about Table 3 labels
Address comment from review 2A about connectionless send

@@
 The API that is exposed to applications using TCP is invisible from the
-network, and can be changed. Relaxing the API to enable out-of-order
+network, and can be changed.  For example, TCP Fast Open \cite{RFC7413} 
+has been implemented by overloading the connectionless \texttt{sendto()} 
+call to trigger an implicit \texttt{connect()} when used on an unconnected
+TCP socket.
+Relaxing the API to enable out-of-order
 delivery of segments is trivial: segments are delivered to the application
@@
 dependencies.
-
-\todo{Review A: Section 3 should also note that connectionless send and
-receive already work on connection-oriented sockets}
 

### 2016-06-22 paper.tex
Expand discussion of codec/cc interactions, to address reviewer 2A comment

@@
 
-\newcommand{\fail}{\tikz\draw[red,fill=red] (0,0) circle (.5ex);}
-\newcommand{\pass}{\tikz\draw[green,fill=green] (0,0) circle (.5ex);}
-\newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) circle (.5ex);}
+\newcommand{\fail}{\tikz\draw[red,fill=red] (0,0) rectangle (1.5ex,1.5ex);}
+\newcommand{\pass}{\tikz\draw[green,fill=green] (0,0) circle (.75ex);}
+\newcommand{\okay}{\tikz\draw[orange,fill=orange] (0,0) -- (0.75ex, 1.5ex) -- (1.5ex,0) -- cycle;}
 
@@
 competes well with standard TCP in many environments, and is seeing 
-commercial deployment. 
-It would also be possible to
-implement alternatives that seek stability, or compatibility with the
-dictates of a video codec, rather than traditional ``TCP Friendly''
-congestion control -- even if implemented within TCP.
-The development of TCP congestion control shows that there might be
-fairness issues as new algorithms are deployed, but the network does not
-prevent the deployment of those algorithms.
+commercial deployment.  The development of TCP congestion control shows
+that there might be fairness issues as new algorithms are deployed, but the
+network does not prevent the deployment of those algorithms.
+
+It would also be possible to implement alternative congestion control
+algorithms that seek stability, or compatibility with the dictates of a
+video codec, rather than traditional ``TCP Friendly'' congestion control,
+even if implemented within TCP.
+To do this effectively might require changes to the interface between the
+application and the TCP stack, even if the on-the-wire format remains the
+same. For example, video applications generate data periodically, and it
+might help the congestion control to know the period, so it can pace out
+data; video traffic is less elastic than many TCP bulk flows, and it might
+be benefical to inform the stack of an upper rate beyond which there is no
+point increasing the congestion window, and a lower rate beyond which the 
+flow cannot proceed; and informing the codec of the RTT and congestion
+window might allow it to better schedule bursts of traffic to match the 
+available capacity. In the interactive video conferencing community, 
+\cite{draft-ietf-rmcat-cc-codec-interactions-02} addresses this issue for
+congestion control over RTP on UDP/IP, but there is no analagous document
+for TCP congestion control interactions as yet.
 
@@
   data, and \fail \thinspace indicates connection failure (none observed)}
-  \todo{the labels need to be distinctive shapes, not just different colours}
   \label{tab:deployability}

### 2016-06-22 paper.tex
In Table 3, note all are examples where UDP fails

@@
 inconsistent retransmissions \cite{mcquistin2016tcp}, testing all
-major UK providers. The results are shown in Table \ref{tab:deployability}.
+major UK providers, with the sender at the University of Glasgow. 
+The results are shown in Table \ref{tab:deployability}.
 We found that 100\% of tested fixed-line networks
@@
   indicates successful delivery, \okay \thinspace indicates delivery of the original
-  data, and \fail \thinspace indicates connection failure (none observed)}
+  data, and \fail \thinspace indicates connection failure (none observed). 
+  We note that the campus firewall near the server blocks UDP traffic, so
+  all are examples where fallback from UDP to TCP is beneficial for
+  real-time traffic.}
   \label{tab:deployability}

### 2016-06-22 paper.tex
Add forward reference to Section 5 for COBS

@@
 network. A framing marker is added to the start and end of each message
-before transmission. An encoding algorithm is used to escape all occurrences
-of the framing marker within the message data. This process does not impact
-on the data that can be sent or received by applications.
+before transmission, and removed on reception, and an encoding algorithm 
+is used to escape all occurrences of the framing marker within the message
+data. This process does not impact on the data that can be sent or received
+by applications. As discussed in Section \ref{sec:realising}, COBS framing
+\cite{CB97COBS} is suitable for this purpose.
 

### 2016-06-22 paper.tex
Remove stray todo
Tweak intro

@@
 congestion control coupled with a reliable, ordered, byte stream API.
-These are proven suitable for many applications, but are inappropriate
-for real-time traffic. While both protocols may be used for real-time
+These have been proven suitable for many applications, but are inappropriate
+for real-time traffic. While both protocols are used for real-time
 applications, neither really provides the right services and API. This
@@
 should be provided by the transport. The increased costs and
-complexity of doing so raise barriers to innovation.
+complexity of doing so make applications less reliable, and raise barriers to innovation.
 
@@
 
-\todo{Review A: Section 4 should have a forward reference to section 5}
 %==================================================================================================

### 2016-06-22 paper.tex
Move Table 1 earlier
Move Table 2 earlier
Clarify discussion of deadlines

@@
 
-% % When compared to conventional real-time systems, networked multimedia
-% % deadlines are neither soft nor hard.
-% Networked multimedia deadlines are unusual when compared to other real-time
-% systems. They are simultaneously flexible and strict: flexible in that the exact
-% value of the deadline is typically not important, provided it is of the right
-% order-of-magnitude for the application, but strict in that any particular
-% deadline provides a hard cut-off, after which the data is useless.
-It is useful to distinguish between networked multimedia deadlines and
-deadlines in conventional real-time systems. A multimedia deadline is neither
-soft, since arrivals played after a deadline contribute to poor experience; nor
-are deadlines hard, since occasional missed messages can be tolerated.
+Networked multimedia deadlines are unusual when compared to other real-time
+systems. They are simultaneously flexible and strict: flexible in that the exact
+value of the deadline is typically not important, provided it is of the right
+order-of-magnitude for the application, but strict in that any particular
+deadline provides a cut-off, after which the data arrives too late to be
+rendered to the user (although, again, it is not entirely useless, since
+it might be used to complete predictive coding state, improving the quality 
+of frames decoded later).
 
@@
 
+\begin{table}
+  \centering
+  \begin{tabular}{ll}
+    \toprule
+      Transport Service     & Requirement \\
+    \midrule
+      Deadlines             & Core        \\
+      Partial reliability   & Core        \\
+      Dependencies          & Core        \\
+      Message-oriented      & Core        \\
+      Sub-streams           & Core        \\
+      Congestion controlled & Core        \\
+      Connection oriented   & Subsidiary  \\
+      Keep-alive            & Subsidiary  \\
+    \bottomrule
+  \end{tabular}
+  \caption{Transport services for real-time multimedia}
+  \label{tab:services}
+\end{table}
+
 %--------------------------------------------------------------------------------------------------
 \subsection{Abstract API}
+
+\begin{table*}[t]
+  \centering
+  \begin{tabularx}{\textwidth}{llll}
+    \toprule
+      Transport Service            & Function & Parameters                                     & Return Value(s) \\
+    \midrule
+                                   & \texttt{socket}
+                                              & \texttt{af} -- Address family                  & Socket descriptor \\
+                                   &          & \texttt{st} -- Socket type & \\
+                                   & \texttt{close}
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+    \midrule
+                                   & \texttt{getsockopt/setsockopt}                                    
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   &          & \texttt{level} -- Protocol level & \\
+                                   &          & \texttt{option} -- Option name & \\
+                                   &          & \texttt{value} -- Option value & \\
+                                   &          & \texttt{len} -- Option length & \\
+    \midrule
+      Connection oriented          & \texttt{bind}
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   &          & \texttt{addr} -- Address to bind to & \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
+                                   & \texttt{listen}
+                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
+                                   & \texttt{accept}
+                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
+                                   &          & \texttt{addr} -- Address of peer & \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
+                                   & \texttt{connect}
+                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
+                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
+    \midrule
+      Deadlines                    & \texttt{set\_po\_delay}
+                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
+    \midrule
+      Message oriented             & \texttt{send\_message}
+                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
+                                   &          & \texttt{buf} -- Message data & \\
+                                   &          & \texttt{len} -- Length of message data & \\
+                                   &          & \texttt{seq\_num} -- Sequence number & \\
+      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
+      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
+      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
+                                   & \texttt{recv\_message}
+                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
+                                   &          & \texttt{buf} -- Buffer for message data        & Substream identifier \\
+                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
+    \bottomrule
+  \end{tabularx}
+  \caption{Outline transport API for real-time applications. Return values
+  shown are for successful calls; in all cases, -1 is returned in the event
+  of an error}
+  \label{tab:api}
+\end{table*}
 
@@
 transmitted until they expire.
-
-\begin{table}
-  \centering
-  \begin{tabular}{ll}
-    \toprule
-      Transport Service     & Requirement \\
-    \midrule
-      Deadlines             & Core        \\
-      Partial reliability   & Core        \\
-      Dependencies          & Core        \\
-      Message-oriented      & Core        \\
-      Sub-streams           & Core        \\
-      Congestion controlled & Core        \\
-      Connection oriented   & Subsidiary  \\
-      Keep-alive            & Subsidiary  \\
-    \bottomrule
-  \end{tabular}
-  \caption{Transport services for real-time multimedia}
-  \label{tab:services}
-\end{table}
-
-\begin{table*}[t]
-  \centering
-  \begin{tabularx}{\textwidth}{llll}
-    \toprule
-      Transport Service            & Function & Parameters                                     & Return Value(s) \\
-    \midrule
-                                   & \texttt{socket}
-                                              & \texttt{af} -- Address family                  & Socket descriptor \\
-                                   &          & \texttt{st} -- Socket type & \\
-                                   & \texttt{close}
-                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
-    \midrule
-                                   & \texttt{getsockopt/setsockopt}                                    
-                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
-                                   &          & \texttt{level} -- Protocol level & \\
-                                   &          & \texttt{option} -- Option name & \\
-                                   &          & \texttt{value} -- Option value & \\
-                                   &          & \texttt{len} -- Option length & \\
-    \midrule
-      Connection oriented          & \texttt{bind}
-                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
-                                   &          & \texttt{addr} -- Address to bind to & \\
-                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
-                                   & \texttt{listen}
-                                              & \texttt{sd} -- Socket descriptor               & 0 (success), -1 (error) \\
-                                   & \texttt{accept}
-                                              & \texttt{sd} -- Listening socket descriptor     & Connection socket descriptor \\
-                                   &          & \texttt{addr} -- Address of peer & \\
-                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
-                                   & \texttt{connect}
-                                              & \texttt{addr} -- Address to connect to         & 0 (success), -1 (error) \\
-                                   &          & \texttt{addrlen} -- Length of \texttt{addr} & \\
-    \midrule
-      Deadlines                    & \texttt{set\_po\_delay}
-                                              & \texttt{delay} -- Playout delay (in ms)        & 0 (success), -1 (error) \\
-    \midrule
-      Message oriented             & \texttt{send\_message}
-                                              & \texttt{sd} -- Socket descriptor               & Number of bytes sent \\
-                                   &          & \texttt{buf} -- Message data & \\
-                                   &          & \texttt{len} -- Length of message data & \\
-                                   &          & \texttt{seq\_num} -- Sequence number & \\
-      Deadlines                    &          & \texttt{deadline} -- Relative deadline of message (in ms) & \\
-      Dependencies                 &          & \texttt{depends\_on} -- \texttt{seq\_num} of dependency & \\
-      Sub-streams                  &          & \texttt{substream} -- Substream identifier & \\
-                                   & \texttt{recv\_message}
-                                              & \texttt{sd} -- Socket descriptor               & Number of bytes received \\
-                                   &          & \texttt{buf} -- Buffer for message data        & Substream identifier \\
-                                   &          & \texttt{len} -- Size of \texttt{buf} & \\
-    \bottomrule
-  \end{tabularx}
-  \caption{Outline transport API for real-time applications. Return values
-  shown are for successful calls; in all cases, -1 is returned in the event
-  of an error}
-  \label{tab:api}
-\end{table*}
 

### 2016-06-22 paper.tex
Further updates to §2.1

@@
 rendered to the user (although, again, it is not entirely useless, since
-it might be used to complete predictive coding state, improving the quality 
+it might be used to complete a predictive coding chain, improving the quality 
 of frames decoded later).
@@
 % , and are directly caused by the lack of partial reliability in TCP.
-For the applications under scrutiny, a missed frame that fails to deliver by its
-deadline is much less disruptive than a stall in play-out.
+For the applications under scrutiny, a missed frame that is not delivered
+by its deadline, while surrounding frames are delivered, is much less
+disruptive than a stall in play-out waiting for repair.
 
 \textbf{Message-oriented Dependencies:}
-The combination of deadlines and partial reliability leads to \emph{dependency
-management} as an important transport service.
+The combination of deadlines and partial reliability makes \emph{dependency
+management} an important transport service.
 % There is no point in sending data that cannot be used because it depends on
@@
 be played out, or it may be needed as part of the application's decoding chain.
-Interdependencies between frames of video exist within a number of codecs,
-including MPEG-1 \cite{le1991mpeg}, and newer codecs such as H.264
-\cite{wiegand2003overview}. As an example, we consider MPEG-1 video:
-I-frames are independent, while P- and B-frames contain only the changes 
-since the previous frame (P) or between frames (B), and so are dependent on
-the successful arrival of other frames. An I-frame may not arrive in time
-to be played out, but may need to be sent to ensure that dependent P- and
-B-frames are useful.
+Interdependencies between frames of video exist within a number of codecs.
+The original MPEG-1 codec \cite{le1991mpeg} divided video frames into three
+types. I-frames were independently encoded, while P- and B-frames contain
+only the changes since the previous frame (P), or between frames (B), and so
+could only be decoded dependent on the successful arrival of other frames.
+Newer codecs, such as H.264 \cite{wiegand2003overview}, use more complex
+and sophisticated versions of the same idea. 
+A consequence is that the sender might know that a frame will not arrive in
+time to be played out, but may need to send it anyway to ensure that the
+receiver can decode any dependent frames sent later in the stream.
 
@@
 We note the importance of congestion control. Historically, real-time many
-applications have required an isochronous channel, that is not subject to
+applications have required an isochronous channel, and have not implemented
 congestion control. This is impractical on the Internet. Further, while
 some applications are non-adaptive or constant bit-rate, an increasing
-number are either or both of adaptive and variable bit-rate. Users would
-be better served by applications that adapt to available bandwidth.
+number are either, or both, of adaptive and variable bit-rate. Users would
+be better served by applications that adapt to available bandwidth. This is
+especially true of mobile applications, where channel capacity can vary
+significantly over time.
 
@@
 session.
-Maintaining per-connection state at the endpoints is helpful for the
-implementation of many forms of congestion control, however. 
+On the other hand, maintaining per-connection state at the endpoints is
+helpful for the implementation of many forms of congestion control.
 Signalling messages indicating start and end of connections can also ease

### 2016-06-22 paper.tex
Editorial fixes throughout

@@
      before it is rendered to the user. The play-out delay is fed back to
-     the sender host, for use as part of the media deadline estimation.
+     the sender, for use as part of the media deadline estimation.
 
@@
     The \texttt{recv\_message()} call returns the sequence number, dependency
-    information, and sub-stream identifier along with any received message,
-    allowing the receiver to direct it to the correct decoding queue.
+    information, and sub-stream identifier along with any received message.
+    This allows the receiver to direct the message to the correct decoding queue.
 
@@
 unlikely to pass the network. We may innovate all we like, provided the
-transport of the future looks like TCP or UDP to middleboxes.\footnote{
+transport of the future looks like TCP or UDP to middleboxes.
   This is inconvenient, certainly, but is not necessarily a bad thing.
@@
   Making changes to this type of infrastructure \emph{should} require
-  careful backwards compatibility \cite{mcquistin:2015:reinterpreting}.
-}
+  careful consideration of backwards compatibility \cite{mcquistin:2015:reinterpreting}.
 
@@
 delay-based algorithm that reduces latency, although it is known to be less
-aggressive than standard TCP, and is prone to starvation; FAST TCP 
+aggressive than standard TCP, and is prone to starvation. FAST TCP 
 \cite{jin:2004:fast-tcp} is a more modern delay-based algorithm that 
@@
 \balance
-\bibliographystyle{abbrv}
+\bibliographystyle{acm}
 \bibliography{paper}

### 2016-06-23 paper.tex
Addresses point 4 in review #25B

@@
     and 4) a sub-stream identifier, used, for example, to differentiate audio,
-    video, sub-title, control, and repair streams.
+    video, sub-title, control, and repair streams. Of this metadata, only the
+    sub-stream identifier is sent on the wire. The sequence number, deadline,
+    and dependency information is used only by the sender to provide the
+    partially reliable service.
 
-    The \texttt{recv\_message()} call returns the sequence number, dependency
-    information, and sub-stream identifier along with any received message.
-    This allows the receiver to direct the message to the correct decoding queue.
+    The \texttt{recv\_message()} call returns the sub-stream identifier and
+    length of the message, along with the received message data. This allows 
+    the receiver to direct the message to the correct decoding queue.
 

### 2016-06-23 paper.tex
Addresses point 2 in review #25B

@@
 
-In order to maintain compatibility with middleboxes, offering partial
-reliability requires using \emph{inconsistent retransmissions}. This means
-that the mapping between message data and TCP sequence numbers is no longer
+Middleboxes in the network have ossified around TCP's reliability
+mechanism: they do not expect gaps in the TCP sequence number space.
+Honda \emph{et al.} \cite{honda:2011:extend-tcp} test the behaviour of
+middleboxes in response to gaps in the TCP sequence number space, showing
+that middleboxes interfere with flows in up to 29\% of tested paths. This
+depends on the mechanism and port number used, with ports used by common
+applications impacted most. To ensure compatibility with these middleboxes,
+offering partial reliability requires using \emph{inconsistent
+retransmissions}. Retransmissions will be triggered as under standard TCP
+to ensure that the sequence number space is filled, but the data in a
+retransmitted segment may not be the same as the original. This means that
+the mapping between message data and TCP sequence numbers is no longer
 static: a given TCP sequence number may be relate to different messages at
@@
 required to allow messages to be uniquely identified.
-\todo{Review B: Inconsistent retransmissions are not well introduced. Should
-be more explicit, and earlier.}
 

### 2016-06-23 paper.tex
Addresses point 3 in review #25B

@@
 layer, this implies a \emph{message oriented} service, that
-maintains application data unit (ADU) boundaries.
+maintains application data unit (ADU) boundaries. Messages are delivered to
+the application in the order they arrive. For example, the combination of
+reliability and in-order delivery in TCP can introduce significant latency:
+incoming segments may be head-of-line blocked waiting for the delivery of
+an earlier segment.
 
@@
 metadata entry at the head of the queue, with its TCP sequence number
-attached.
+attached. This means that segments are delivered in the order that they
+arrive, removing the latency associated with head-of-line blocking in TCP
+\cite{mcquistin2016tcp}.
 
@@
 for complete messages (i.e., data between two zero bytes), which are
-decoded and passed to the application. Any message fragments are buffered,
-alongside their TCP sequence number, awaiting the arrival of the remainder
-of the message. Once the message has been reassembled, it is decoded, and
-delivered to the application.
-\todo{Review B: Head-of-line blocking problem not introduced, so not clear
-why this mechanism is needed.}
+decoded and passed to the application. While the segments are sent
+containing only one message, these may be resegmented or coalesced in the
+network. A segment may arrive containing \emph{fragments} of message
+data. These fragments are buffered, alongside their TCP sequence number, 
+awaiting the arrival of the remainder of the message. Once the message has
+been reassembled, it is decoded, and delivered to the application.
 % TCP Hollywood inherits standard TCP's congestion control mechanism.

### 2016-06-23 paper.tex
Corrections

@@
 maintains application data unit (ADU) boundaries. Messages are delivered to
-the application in the order they arrive. For example, the combination of
-reliability and in-order delivery in TCP can introduce significant latency:
-incoming segments may be head-of-line blocked waiting for the delivery of
-an earlier segment.
+the application in the order they arrive. As seen in TCP, in-order delivery
+can introduce significant latency: incoming segments may be head-of-line
+blocked waiting for the delivery of an earlier segment.
 
@@
 for complete messages (i.e., data between two zero bytes), which are
-decoded and passed to the application. While the segments are sent
+decoded and passed to the application. While segments are sent
 containing only one message, these may be resegmented or coalesced in the

### 2016-06-23 paper.tex
Back to abbrv bib style, since acm throws TeX warnings
Scale tables and figures to \linewidth
Tidy markup

@@
 
-\newcommand{\ie}{{i.e.,}\xspace}
 \newcommand*\rot{\rotatebox{90}}
@@
 such as telephony, video conferencing, or telepresence, require low end-to-end
-latency. Their deadlines for presenting the media, \ie playing the audio and
+latency. Their deadlines for presenting the media, i.e., playing the audio and
 displaying the video frame, range from tens to a few hundred milliseconds.
@@
   \centering
-  \begin{tabularx}{\textwidth}{llll}
+  \begin{tabularx}{\linewidth}{llll}
     \toprule
@@
  \centering
- \includegraphics[scale=0.45]{figures/tcp-hollywood-fixed.pdf}
+  \includegraphics[width=\linewidth]{figures/tcp-hollywood-fixed.pdf}
   \caption{TCP Hollywood architecture}
@@
 \balance
-\bibliographystyle{acm}
+\bibliographystyle{abbrv}
 \bibliography{paper}

### 2016-06-23 paper.tex
Minor editorial fixes

@@
 These issues are slowly changing, as UDP-based applications penetrate the
-enterprise consciousness, but not clear that UDP is universally available
+enterprise consciousness, but UDP is not universally available
 (Google report 90-95\% of endpoints are reachable with QUIC
@@
 
-The API that is exposed to applications using TCP is invisible from the
+The API that is exposed to applications using TCP is invisible to the
 network, and can be changed.  For example, TCP Fast Open \cite{RFC7413} 
@@
 mechanism: they do not expect gaps in the TCP sequence number space.
-Honda \emph{et al.} \cite{honda:2011:extend-tcp} test the behaviour of
+Honda \emph{et al.}\ \cite{honda:2011:extend-tcp} test the behaviour of
 middleboxes in response to gaps in the TCP sequence number space, showing
@@
 different times. Therefore, an application-level sequence number is
-required to allow messages to be uniquely identified.
+required to allow messages to be uniquely identified (multipath TCP has
+a similar requirement).
 
@@
 on the network that are performing payload inspection. These middleboxes
-may interpret this behaviour relating to an attack. For example, a
+may interpret this behaviour as relating to an attack. For example, a
 man-on-the-side attack exhibits similar behaviour, where a malicious host
@@
 We conducted further deployment experiments using
-inconsistent retransmissions \cite{mcquistin2016tcp}, testing all
-major UK providers, with the sender at the University of Glasgow. 
+inconsistent retransmissions, testing all
+major UK providers, with the sender at the University of Glasgow \cite{mcquistin2016tcp}. 
 The results are shown in Table \ref{tab:deployability}.
 We found that 100\% of tested fixed-line networks
-delivered inconsistent retransmissions successfully. However, the delivery
+delivered inconsistent retransmissions successfully. However, delivery
 of the original segment is common on cellular networks, with only 25\% of
@@
 
-Use of inconsistent retransmissions may interact negatively with
-caching and re-segmenting middleboxes, resulting in the corruption of
-messages between sender and receiver. A message may be formed from the
-original message, and an inconsistent retransmission, given how the
-mechanism uses the TCP sequence space. To protect against this, a checksum
+Use of inconsistent retransmissions can interact negatively with
+middleboxes that cache and re-segment TCP streams, resulting in the corruption of
+messages between sender and receiver. The result can be a message formed
+from some combination of the
+original message and an inconsistent retransmission.
+To protect against this, a checksum
 must be attached to each message, to allow the receiver to verify its
@@
 While further measurement studies are required to confirm the ability to
-deploy wire-visible changes to TCP (such as inconsistent retransmissions),
+deploy wire-visible changes to TCP (such as inconsistent retransmissions)
+in the wider Internet,
 we have shown that we can provide all of the transport services needed by
